### 2024-03-31 paper/paper.qmd
Updated paper.qmd with graphs

---
title: "My title"
subtitle: "My subtitle if needed"
author: Emily Su
thanks: "Code and data are available at: https://github.com/moonsdust/top-songs."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(ggcorrplot)
library(ggplot2)
library(knitr)
library(DiagrammeR)

# To be removed because this is from example
library(palmerpenguins)
analysis_data <- read_csv('../data/analysis_data/analysis_data.csv')

# Read in data 
playlist_analysis_data <- read_parquet("../data/analysis_data/playlists_analysis_data.parquet")
```

# Introduction

The estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.

# Data {#sec-data}

## Variables of Interest

## Data Source and Measurements

## Methodology

# Model {#sec-model}

```{r}
#| label: fig-causal-model
#| fig-cap: Causal relationship between song characteristics and hit song (before 2020 or during and after 2020)
#| echo: false
#| warning: false
#| message: false

# Referenced https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/
grViz("
digraph {
  graph []
  node [shape = plaintext]
    A [label = 'Hit song before 2020']
    B [label = 'Modality']
    C [label = 'Loudness']
    D [label = 'Duration']
    E [label = 'Tempo']
  edge []
    B->A
    C->A
    D->A
    E->A
  { rank = same; }
}
")
```

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic)

### Scale and Modality
```{r}
#| label: fig-key-mode-bar
#| fig-cap: Proportion of songs in different scales before 2020 versus during and after 2020
#| echo: false
#| warning: false
#| message: false

proportion_of_each_key_mode <- 
  playlist_analysis_data |>
  # Group by if song was a hit before or during and after the pandemic
  group_by(period) |>
  # Count each scale
  count(key_mode) |>
  # rename count column to num_of_key_mode
  rename(
    'num_of_key_mode' = n
  ) |>
  mutate( 
    proportion_key_mode =  round(num_of_key_mode / sum(num_of_key_mode), 2)
  )

proportion_of_each_key_mode |> 
  ggplot(mapping = aes(x = key_mode, y = proportion_key_mode)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_bar(stat="identity", binwidth = 3) +
  theme_minimal() +
  labs(x = "Scale",
       y = "Proportion") + 
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
#| label: fig-mode-bar
#| fig-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
#| echo: false
#| warning: false
#| message: false

proportion_of_each_mode <- 
  playlist_analysis_data |>
  # Group by if song was a hit before or during and after the pandemic
  group_by(period) |>
  # Count each scale
  count(mode_name) |>
  # rename count column to num_of_key_mode
  rename(
    'num_of_mode_name' = n
  ) |>
  mutate( 
    proportion_mode_name =  round(num_of_mode_name / sum(num_of_mode_name), 2) 
  )

proportion_of_each_mode |> 
  ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_bar(stat = "identity", binwidth = 3) +
  theme_minimal() +
  labs(x = "Modality",
       y = "Proportion") + 
  theme(axis.text.x = element_text(angle = 90))
```
```{r}
#| label: tbl-mode-table
#| tbl-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
#| echo: false
#| warning: false
#| message: false

proportion_of_each_mode |> 
  rename( 
    "Period" = period,
    "Modality" = mode_name,
    "Count of each mode" = num_of_mode_name,
    "Proportion of each mode" = proportion_mode_name
    ) |>
  kable()
```

### Track Duration Over the Years

```{r}
#| label: fig-track-duration-scatterplot
#| fig-cap: Relationship between the hit year and track duration of a song 
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = track_duration_ms)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Duration of song (in ms)")
```

```{r}
#| label: fig-track-duration-hist
#| fig-cap: Distribution of track duration (in ms) before 2020 versus during and after 2020  
#| echo: false
#| warning: false
#| message: false
#| 
playlist_analysis_data |> 
  ggplot(mapping = aes(x = track_duration_ms)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 7000) +
  theme_minimal() +
  labs(x = "Duration of song (in ms)",
       y = "Count")
```

```{r}
#| label: fig-track-duration-boxplot
#| fig-cap: Track duration (in ms) before 2020 versus during and after 2020  
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(mapping = aes(x = track_duration_ms)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Duration of song (in ms)",
       y = "Count")
```

```{r}
#| label: tbl-track-duration-stats-before
#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) before 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose track_duration_ms
  select(track_duration_ms) |>
  # Rename columns
  rename (
    "Duration of song (in ms)"= track_duration_ms
  ) |>
  summary() |>
  kable()
```
```{r}
#| label: tbl-track-duration-stats-during-after
#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for During and After Pandemic
  filter(period == "During and After Pandemic") |>
  # Choose track_duration_ms
  select(track_duration_ms) |>
  # Rename columns
  rename (
    "Duration of song (in ms)"= track_duration_ms
  ) |>
  summary() |>
  kable()
```

### Loudness

```{r}
#| label: fig-loudness-scatterplot
#| fig-cap: Relationship between the hit year and loudness of a song 
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = loudness)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Loudness of song (in dB)")
```
```{r}
#| label: fig-loudness-hist
#| fig-cap: Distribution of loudness (in dB) before 2020 versus during and after 2020  
#| echo: false
#| warning: false
#| message: false
#| 
playlist_analysis_data |> 
  ggplot(mapping = aes(x = loudness)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 0.3) +
  theme_minimal() +
  labs(x = "Loudness of song (in dB)",
       y = "Count")
```

```{r}
#| label: tbl-loudness-stats-before
#| tbl-cap: Minimum, quartiles, median, and maximum of loudness of song (in dB) before 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose loudness
  select(loudness) |>
  # Rename columns
  rename (
    "Loudness of song (in dB)"= loudness
  ) |>
  summary() |>
  kable()
```

```{r}
#| label: tbl-loudness-stats-during-after
#| tbl-cap: Minimum, quartiles, median, and maximum of loudness of song (in dB) during and after 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "During and After Pandemic"
  filter(period == "During and After Pandemic") |>
  # Choose loudness
  select(loudness) |>
  # Rename columns
  rename (
    "Loudness of song (in dB)"= loudness
  ) |>
  summary() |>
  kable()
```

### Tempo

```{r}
#| label: fig-tempo-scatterplot
#| fig-cap: Relationship between the hit year and tempo of a song 
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = tempo)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Tempo of song (BPM)")
```

```{r}
#| label: fig-tempo-hist
#| fig-cap: Distribution of tempo (BPM) before 2020 versus during and after 2020  
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(mapping = aes(x = tempo)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 3) +
  theme_minimal() +
  labs(x = "Tempo of song (BPM)",
       y = "Count")
```

```{r}
#| label: tbl-tempo-stats-before
#| tbl-cap: Minimum, quartiles, median, and maximum of tempo of a song (BPM) before 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose tempo
  select(tempo) |>
  # Rename columns
  rename (
    "Tempo of song (BPM)"= tempo
  ) |>
  summary() |>
  kable()
```

```{r}
#| label: tbl-tempo-stats-during-after
#| tbl-cap: Minimum, quartiles, median, and maximum of tempo of a song (BPM) during and after 2020
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "During and After Pandemic"
  filter(period == "During and After Pandemic") |>
  # Choose tempo
  select(tempo) |>
  # Rename columns
  rename (
    "Tempo of song (BPM)"= tempo
  ) |>
  summary() |>
  kable()
```

## Model Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = "../models/first_model.rds")
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Additional Figures and Tables

```{r}
#| label: fig-correlation-matrix
#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 single and if song was a hit before 2020/pandemic
#| echo: false
#| warning: false
#| message: false

# Referenced https://rpubs.com/Alema/1000474
numeric_playlist_analysis_data <- 
  playlist_analysis_data |>
  select(-c(mode_name, key_mode, period, hit_year))

ggcorrplot(cor(numeric_playlist_analysis_data), 
           hc.order = TRUE,
           outline.color = "white", lab = TRUE, legend.title = "Correlation") 
```

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-31 scripts/00-install_packages.R
Added rsvg package into install packages script

@@ -20,5 +20,6 @@ install.packages("ggcorrplot") # For correlation matrix
 install.packages("testthat") # To test
 # For models
 install.packages("DiagrammeR")
+install.packages("rsvg")
 install.packages("rstanarm")
 install.packages("modelsummary")

### 2024-03-31 scripts/00-install_packages.R
Updated install packages script

@@ -21,5 +21,9 @@ install.packages("testthat") # To test
 # For models
 install.packages("DiagrammeR")
 install.packages("rsvg")
+install.packages("magrittr")
+install.packages("DiagrammeRsvg")
+install.packages("png")
+
 install.packages("rstanarm")
 install.packages("modelsummary")

### 2024-03-31 scripts/04-model.R
Finished model script

@@ -1,36 +1,38 @@
 #### Preamble ####
-# Purpose: Models information about songs from Spotify playlists
+# Purpose: Models information about songs from Billboard Year-End Hot 100
 # Author: Emily Su
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R, 01-download_data.R, 02-data_cleaning.R has been ran to install necessary packages.
+# Pre-requisites: 00-install_packages.R, 01-download_data.R, 02-data_cleaning.R, and
+# 03-test_data.R has been ran to install necessary packages, download raw data, and 
+# clean the data, and test the data, respectively.
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
+
+set.seed(646)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/analysis_data/playlists_analysis_data.parquet")
 
 ### Model data ####
-first_model <-
+
+# Logistic regression model for hit song before 2020/pandemic
+hit_song_model <-
   stan_glm(
-    formula = flying_time ~ length + width,
+    formula = before_pandemic ~ track_duration_ms + loudness + tempo + (major + minor),
     data = analysis_data,
-    family = gaussian(),
+    family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+    seed = 646
   )
 
 
 #### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
-
+saveRDS(hit_song_model, file = "models/hit_song_model.rds")

### 2024-03-31 scripts/04-model.R
Updated hit song model and model script

@@ -33,6 +33,7 @@ hit_song_model <-
     seed = 646
   )
 
+prior_summary(hit_song_model)
 
 #### Save model ####
 saveRDS(hit_song_model, file = "models/hit_song_model.rds")

### 2024-03-31 scripts/04-model.R
Updated model formula

@@ -24,7 +24,7 @@ analysis_data <- read_parquet("data/analysis_data/playlists_analysis_data.parque
 # Logistic regression model for hit song before 2020/pandemic
 hit_song_model <-
   stan_glm(
-    formula = before_pandemic ~ track_duration_ms + loudness + tempo + (major + minor),
+    formula = before_pandemic ~ track_duration_ms + loudness + tempo + mode_name,
     data = analysis_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-31 paper/paper.qmd
Updated referenced and paper.qmd

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Characteristics of Top Songs Has Changed from Pandemic Brain"
+subtitle: "An analysis of songs on Billboard's Year-End Hot 100 list (2014 to 2023)"
 author: Emily Su
@@
 library(knitr)
+
+# Packages related to causal model created
 library(DiagrammeR)
+library(rsvg)
+library(magrittr)
+library(DiagrammeRsvg)
+library(png)
 
@@
 ## Methodology
+The datasets used in this paper were retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot],
+testthat [@citetestthat].
+The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
 
@@
 # Referenced https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/
-grViz("
-digraph {
+# https://stackoverflow.com/questions/42737645/how-to-export-images-of-diagrammer-in-r
+causal_model <- "digraph {
   graph []
   node [shape = plaintext]
-    A [label = 'Hit song before 2020']
+    A [label = 'Hit song before 2020 or during and after 2020']
     B [label = 'Modality']
@@
   { rank = same; }
-}
-")
+}"
+
+exporting_image <- grViz(causal_model) |>
+  export_svg() |>
+  charToRaw() |>
+  rsvg_png("casual_model.png", width = 1000)
+
+# NOTE: Had to save diagram and read it in as a image since grViz can be only render in HTML and 
+# not in PDFs 
+# Referenced https://stackoverflow.com/questions/23861000/displaying-images-in-r-in-version-3-1-0
+the_image <- readPNG("casual_model.png")
+plot.new()
+rasterImage(the_image, 0, 0, 0.9, 0.9)
 ```
@@
 
-## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic)
+## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic) of songs from the Billboard Year-End Hot 100 singles
 
@@
 ```
+\newpage
 
-### Track Duration Over the Years
+### Track Duration
 
@@
 
+\newpage
+
 ### Loudness
@@
 ```
+
+\newpage
 
@@
 
+\newpage
+
 ## Model Results
@@
 #| label: fig-correlation-matrix
-#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 single and if song was a hit before 2020/pandemic
+#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 single and if song was a hit before 2020/the pandemic
 #| echo: false

### 2024-04-01 paper/paper.qmd
Updated causal model, references, and paper.qmd

@@
 
+1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper.
+
+Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the COVID-19 pandemic and the lockdowns that came with it [@citecoping].
+This finding is similar to how music therapy works.
+As defined by Canadian Association of Music Therapists, music therapy is the use of music to support an individual's health, social development, and well-being and some of the techniques used includes listening to music [@citemusictherapy].
+
+Hurwitz and Krumhansl conducted a study into how people's listening habits over the course of the pandemic and they found that the songs they listened to often is linked to an emotion such as sadness and/or invokes memories from the past [@citeshiftingniches].
+
+During the COVID-19 pandemic, TikTok grew in popularity alongside shortform videos for different types of content.
+With this, viral songs on TikTok were met with thousands of listens on streaming platforms such as Spotify and sometimes go onto being on lists like the Billboard Year-End Hot 100.
+
+Music reflects the climate of society.
+
+However this raises the following question: how are top songs prior to 2020 different from the top songs during and after 2020?
+To investigate this question, we looked at their music characteristics such as its modality (major or minor key), track duration, loudness, and tempo.
+
 The estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.
+
+Importance: 
+- Being cognizant of music that is popular can help 
+- Can help us figure out the overall feelings of society in today's climate after the pandemic.
+- Results from this paper can help improve music therapy.
 
@@
   graph []
+  # Nodes 
   node [shape = plaintext]
@@
     E [label = 'Tempo']
+    F [label = 'Lyrics (Confounder)']
+  # Edges for non-confounding variables
   edge []
@@
 { rank = same; }
+  # Edges for confounding variable
+  edge [style = dashed]
+    F -> B
+    F -> C
+    F -> D
+    F -> E
+    F -> A
 }"
@@
 plot.new()
-rasterImage(the_image, 0, 0, 0.9, 0.9)
+rasterImage(the_image, 0, 0, 1, 1)
 ```

### 2024-04-01 paper/paper.qmd
Updated paper.qmd's causal model size

@@
 # Introduction
-
-1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper.
 
@@
 plot.new()
-rasterImage(the_image, 0, 0, 1, 1)
+rasterImage(the_image, 0, 0, 0.7, 1)
 ```
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm. We use the default priors from `rstanarm`.
 

### 2024-04-01 paper/paper.qmd
Updated causal model size

@@
 plot.new()
-rasterImage(the_image, 0, 0, 0.7, 1)
+rasterImage(the_image, 0, 0, 0.9, 1.13)
 ```

### 2024-04-01 paper/paper.qmd
Almost done preliminary introduction

--- 2024-04-01 paper/paper.qmd
@@
 
-Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the COVID-19 pandemic and the lockdowns that came with it [@citecoping].
-This finding is similar to how music therapy works.
+During the COVID-19 pandemic, viral songs on social media sites such as TikTok were met with thousands of listens on streaming platforms such as Spotify and sometimes go onto being on lists like the Billboard Year-End Hot 100.
+These songs often accompany thousands of short clips of people talking about their life, challenges, etc.
+Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
+We can see this translated through the shortform videos where users don't talk and instead have the song play out to convey how they are feeling or thinking.
+Ghaffari et al's findings is similar to how music therapy works.
 As defined by Canadian Association of Music Therapists, music therapy is the use of music to support an individual's health, social development, and well-being and some of the techniques used includes listening to music [@citemusictherapy].
+Hurwitz and Krumhansl conducted a study into how people's listening habits over the course of the pandemic and they found that the songs they listened to often is linked to an emotion such as sadness and/or invokes memories from the past [@citeshiftingniches].
+However this raises the following question, which we explored in our analysis: how are top songs prior to 2020 or the start of the pandemic different from the top songs during and after 2020?
 
-Hurwitz and Krumhansl conducted a study into how people's listening habits over the course of the pandemic and they found that the songs they listened to often is linked to an emotion such as sadness and/or invokes memories from the past [@citeshiftingniches].
+In this paper, to investigate patterns and trends in music, we analyzed data from Billboard Year-End Hot 100 singles list from 2014 to 2023 on music characteristics such as its modality (major or minor key), track duration, loudness, and tempo.
+Our estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.
+The list allowed us to gauge how preference in music characteristics has changed overall during and after the pandemic.
+Current studies gives us a sense on how currently people are feeling based on surveys done.
+However, there is currently a lack of understanding and focus on the difference in musical characteristics of popular songs with the general population before the pandemic and during and after it.
+We found that ......[add in stuff].
+Music can reflect the climate of society and being cognizant of the trends in music characteristics found in hit songs can help provide insight on the overall feelings of society in today's climate during and after the pandemic as well as before it.
+This can aid in improving therapeutic treatments such as music therapy.
 
-During the COVID-19 pandemic, TikTok grew in popularity alongside shortform videos for different types of content.
-With this, viral songs on TikTok were met with thousands of listens on streaming platforms such as Spotify and sometimes go onto being on lists like the Billboard Year-End Hot 100.
-
-Music reflects the climate of society.
-
-However this raises the following question: how are top songs prior to 2020 different from the top songs during and after 2020?
-To investigate this question, we looked at their music characteristics such as its modality (major or minor key), track duration, loudness, and tempo.
-
-The estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.
-
-Importance: 
-- Being cognizant of music that is popular can help 
-- Can help us figure out the overall feelings of society in today's climate after the pandemic.
-- Results from this paper can help improve music therapy.
+In the rest of this paper, the data section [@sec-data] will cover the dataset used, how it was obtained, define our variables of interests that are used by our tables and graphs, and briefly explain the data cleaning process.
+In the model section [@sec-model], it will explain go over our proposed causal model, which would explain potential relationships in our data, the setup of our model to understand these relationships, and justification for our model.
+In the results section [@sec-results], we will reveal tables and graphs made on our datasets, explain what they show, and show our results from our model.
+In the discussion section [@sec-discussion], we will connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
+Finally, the appendix section [@sec-appendix] will extra tables and graphs from our results as well, additional information about the model, and a link to a Shiny app featuring an interactive graph of some of the results.
 
@@
 
-```{r}
+```{r, out.width="50%"}
 #| label: fig-causal-model
@@
 plot.new()
-rasterImage(the_image, 0, 0, 0.7, 1)
+rasterImage(the_image, 0, 0, 0.8, 1.13)
 ```
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
@@
 
-# Model details {#sec-model-details}
+## Model details {#sec-model-details}
 
-## Posterior predictive check
+### Posterior predictive check
 
@@
 
-## Diagnostics
+### Diagnostics
 

### 2024-04-02 paper/paper.qmd
Updated model setup, finished abstract, and initial introduction (excluding model results)

--- 2024-04-01 paper/paper.qmd
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Music often reflects the current climate of society and there is a growing interest in how hit songs as seen on Billboard's Year-End Hot 100 singles has changed after the COVID-19 pandemic.
+This paper looks at music characteristics such as tempo, song duration, loudness, and modality of songs from Billboard's Year-End Hot 100 singles list from 2014 to 2023 to reveal patterns and relationships to explain the difference between top songs before 2020 and 2020 onwards.
+The results show that hit songs from 2020 onwards had become on average shorter, quieter, but slightly faster with the melody of songs being in a major key.
+These results can support the evaluation of the current emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
 format:
@@
 
-Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the COVID-19 pandemic and the lockdowns that came with it [@citecoping].
-This finding is similar to how music therapy works.
-As defined by Canadian Association of Music Therapists, music therapy is the use of music to support an individual's health, social development, and well-being and some of the techniques used includes listening to music [@citemusictherapy].
+During the COVID-19 pandemic, viral songs on social media sites such as TikTok were met with thousands of listens on streaming platforms such as Spotify and at times go onto being on lists like the Billboard Year-End Hot 100.
+These songs often accompany thousands of short clips of people talking about their lives, challenges, etc.
+Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
+We can see this translated through the short-form videos where users don't talk and instead have the song play out to convey how they are feeling or thinking.
+Ghaffari et al's findings are similar to how music therapy works.
+As defined by the Canadian Association of Music Therapists, music therapy is the use of music to support an individual's health, social development, and well-being and some of the techniques used include listening to music [@citemusictherapy].
+Hurwitz and Krumhansl conducted a study into how people's listening habits throughout the pandemic and they found that the songs they listened to often were linked to an emotion such as sadness and/or invoking memories from the past [@citeshiftingniches].
+However, this raises the following question, which we explored in our analysis: how are top songs prior to 2020 or the start of the pandemic different from the top songs during and after 2020?
 
-Hurwitz and Krumhansl conducted a study into how people's listening habits over the course of the pandemic and they found that the songs they listened to often is linked to an emotion such as sadness and/or invokes memories from the past [@citeshiftingniches].
+In this paper, to investigate patterns and trends in music, we analyzed data from the Billboard Year-End Hot 100 singles list from 2014 to 2023 on music characteristics such as its modality (major or minor key), track duration, loudness, and tempo.
+Our estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.
+The list allowed us to gauge how preference in music characteristics has changed overall during and after the pandemic.
+Current studies give us a sense of how currently people are feeling based on surveys done.
+However, there is currently a lack of understanding and focus on the difference in musical characteristics of popular songs with the general population before the pandemic and during and after it.
+In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020.
+Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight on the overall feelings of different populations during and after the pandemic as well as before it.
+This can aid in improving treatments such as music therapy.
 
-During the COVID-19 pandemic, TikTok grew in popularity alongside shortform videos for different types of content.
-With this, viral songs on TikTok were met with thousands of listens on streaming platforms such as Spotify and sometimes go onto being on lists like the Billboard Year-End Hot 100.
-
-Music reflects the climate of society.
-
-However this raises the following question: how are top songs prior to 2020 different from the top songs during and after 2020?
-To investigate this question, we looked at their music characteristics such as its modality (major or minor key), track duration, loudness, and tempo.
-
-The estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020.
-
-Importance: 
-- Being cognizant of music that is popular can help 
-- Can help us figure out the overall feelings of society in today's climate after the pandemic.
-- Results from this paper can help improve music therapy.
+In the rest of this paper, the data section (@sec-data) will cover the dataset used, how it was obtained, define the variables of interest that are used by our tables and graphs, and briefly explain the data cleaning process.
+The model section (@sec-model) will explain our proposed causal model, which would explain potential relationships in our data, the setup of our model to understand these relationships, and justification for our model.
+In the results section (@sec-results), we will reveal tables and graphs made on our datasets, explain what they show, and show our results from our model.
+In the discussion section (@sec-discussion), we will connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
+Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring an interactive graph of some of the results.
 
@@
 ## Methodology
-The datasets used in this paper were retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot],
-testthat [@citetestthat].
+The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], testthat [@citetestthat].
 The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
@@
 
+We are interested in investigating if a song is likely to be a top song prior to 2020 or not based on what we know about the duration, loudness, tempo, and modality of a top song.
+
 ## Model set-up
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+We will define our logistic regression model as follows.
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{duration}_i + \beta_2 \times \mbox{loudness}_i + \beta_3 \times \mbox{tempo}_i + \beta_4 \times \mbox{modality}_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @citerstanarm. We use the default priors from `rstanarm`.
+We define $y_i$ to be a top song, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
+$\pi_i$ is the probability that a top song $i$ is a top song prior to 2020.
+Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB).
+Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
+
+We ran the model in R [@citeR] using the rstanarm package of @citerstanarm and we used the default priors from rstanarm.
 

### 2024-04-02 paper/paper.qmd
Removed more used info and files

@@
 The results show that hit songs from 2020 onwards had become on average shorter, quieter, but slightly faster with the melody of songs being in a major key.
-These results can support the evaluation of the current emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
+These results can support the evaluation of the emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
 format:
@@
 library(png)
-
-# To be removed because this is from example
-library(palmerpenguins)
-analysis_data <- read_csv('../data/analysis_data/analysis_data.csv')
 
@@
 # Data {#sec-data}
+[To Do]
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+[To Do]
 
@@
 
-We ran the model in R [@citeR] using the rstanarm package of @citerstanarm and we used the default priors from rstanarm.
+We ran the model using the rstanarm package of @citerstanarm and we used the default priors from rstanarm.
 
 ### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+[To Do]
 
 # Results {#sec-results}
+[To Finish]
 
@@
 ## Model Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = "../models/first_model.rds")
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
+[To Do]
 
 # Discussion {#sec-discussion}
-
+[To Do]
 ## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 ## Model details {#sec-model-details}
+[To Do]
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ### Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-04-05 paper/paper.qmd
Updated paper.qmd and pdf
Updated paper

@@
   charToRaw() |>
-  rsvg_png("casual_model.png", width = 1000)
+  rsvg_png("../other/sketches/graph_2_casual_model.png", width = 1000)
 
@@
 # Referenced https://stackoverflow.com/questions/23861000/displaying-images-in-r-in-version-3-1-0
-the_image <- readPNG("casual_model.png")
+the_image <- readPNG("../other/sketches/graph_2_casual_model.png")
 plot.new()
@@
 # Discussion {#sec-discussion}
+
 [To Do]
-## First discussion point {#sec-first-point}
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
 

### 2024-04-13 scripts/04-model.R
Updated model script

@@ -29,7 +29,6 @@ hit_song_model <-
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
     seed = 646
   )
 

### 2024-04-14 paper/paper.qmd
Updated paper.qmd to show dataset preview
Removed new page command in results section
Added table and graph related to model result

@@
 library(knitr)
+library(modelsummary)
 
@@
 playlist_analysis_data <- read_parquet("../data/analysis_data/playlists_analysis_data.parquet")
+
+hit_song_model_results <- readRDS("../models/hit_song_model.rds")
+
 ```
@@
 In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020.
+[TODO: ADD RESULTS FROM MODEL].
 Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight on the overall feelings of different populations during and after the pandemic as well as before it.
@@
 ## Data Source and Measurements
+```{r}
+#| label: tbl-dataset-preview
+#| tbl-cap: Preview of dataset on Billboard Year-End Hot 100 singles from 2014 to 2023 provided by Spotify
+#| echo: false
+#| warning: false
+#| message: false
+
+playlist_analysis_data |> 
+  head() |>
+  kable()
+```
 
@@
 ```
-\newpage
 
@@
 
-\newpage
-
 ### Loudness
@@
 ```
-
-\newpage
 
@@
 
-\newpage
-
 ## Model Results
 [To Do]
+
+```{r}
+#| label: tbl-model-result
+#| tbl-cap: How likely a hit song is a hit before 2020 / the pandemic based on its tempo, song duration, loudness, and modality
+#| echo: false
+#| warning: false
+#| message: false
+
+modelsummary(
+  list("A hit song before 2020" = hit_song_model_results),
+  statistic = "mad"
+)
+```
+```{r}
+#| label: fig-model-result-ci
+#| fig-cap: The credible intervals for predictors of a hit song before 2020
+#| echo: false
+#| warning: false
+#| message: false
+
+modelplot(hit_song_model_results, conf_level = 0.9) +
+  labs(x = "90 per cent credibility interval")
+```
 

### 2024-04-14 paper/paper.qmd
Updated formatting of paper and added more graphs for the model

@@
 library(modelsummary)
+library(rstanarm)
 
@@
 playlist_analysis_data |> 
+  select(hit_year, track_duration_ms, tempo, loudness, mode_name, key_mode, period) |>
   head() |>
@@
 ```
+Not seen in @tbl-dataset-preview, before_pandemic, major, and minor are numerical versions of the mode_name and period columns.
 
@@
 ### Scale and Modality
+
+```{r}
+#| label: fig-mode-bar
+#| fig-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
+#| echo: false
+#| warning: false
+#| message: false
+
+proportion_of_each_mode <- 
+  playlist_analysis_data |>
+  # Group by if song was a hit before or during and after the pandemic
+  group_by(period) |>
+  # Count each scale
+  count(mode_name) |>
+  # rename count column to num_of_key_mode
+  rename(
+    'num_of_mode_name' = n
+  ) |>
+  mutate( 
+    proportion_mode_name =  round(num_of_mode_name / sum(num_of_mode_name), 2) 
+  )
+
+proportion_of_each_mode |> 
+  ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
+  facet_wrap(facets = vars(period), dir = "v") + 
+  geom_bar(stat = "identity", binwidth = 3) +
+  theme_minimal() +
+  labs(x = "Modality",
+       y = "Proportion") + 
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+```{r}
+#| label: tbl-mode-table
+#| tbl-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
+#| echo: false
+#| warning: false
+#| message: false
+
+proportion_of_each_mode |> 
+  rename( 
+    "Period" = period,
+    "Modality" = mode_name,
+    "Count of each mode" = num_of_mode_name,
+    "Proportion of each mode" = proportion_mode_name
+    ) |>
+  kable()
+```
+
 ```{r}
@@
 
-```{r}
-#| label: fig-mode-bar
-#| fig-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
-#| echo: false
-#| warning: false
-#| message: false
-
-proportion_of_each_mode <- 
-  playlist_analysis_data |>
-  # Group by if song was a hit before or during and after the pandemic
-  group_by(period) |>
-  # Count each scale
-  count(mode_name) |>
-  # rename count column to num_of_key_mode
-  rename(
-    'num_of_mode_name' = n
-  ) |>
-  mutate( 
-    proportion_mode_name =  round(num_of_mode_name / sum(num_of_mode_name), 2) 
-  )
-
-proportion_of_each_mode |> 
-  ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
-  facet_wrap(facets = vars(period), dir = "v") + 
-  geom_bar(stat = "identity", binwidth = 3) +
-  theme_minimal() +
-  labs(x = "Modality",
-       y = "Proportion") + 
-  theme(axis.text.x = element_text(angle = 90))
-```
-```{r}
-#| label: tbl-mode-table
-#| tbl-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
-#| echo: false
-#| warning: false
-#| message: false
-
-proportion_of_each_mode |> 
-  rename( 
-    "Period" = period,
-    "Modality" = mode_name,
-    "Count of each mode" = num_of_mode_name,
-    "Proportion of each mode" = proportion_mode_name
-    ) |>
-  kable()
-```
+\newpage
 
@@
 ```
+
 ```{r}
@@
 ```
+\newpage
 
@@
 
-## Model details {#sec-model-details}
-[To Do]
-
-### Posterior predictive check
-
-### Diagnostics
-
 \newpage
 
+## Posterior predictive check for model {#sec-model-details-posterior-check}
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-posteriorpredictive
+#| fig-cap: "How the data impacts how the model fits"
+#| fig-subcap: ["Posterior prediction check for the model"]
+#| layout-ncol: 1
+
+rstanarm::pp_check(hit_song_model_results) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+```
+\newpage
+
+## Diagnostics for model {#sec-model-details-diagnostics}
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheck
+#| fig-cap: "Checking the convergence of the MCMC algorithm for hit song model"
+#| fig-subcap: ["Trace plot for model", "Rhat plot for model"]
+#| layout-ncol: 2
+
+plot(hit_song_model_results, "trace")
+
+plot(hit_song_model_results, "rhat") + theme_minimal()
+```
+
 # References

### 2024-04-16 paper/paper.qmd
removed numbering from equation
Finished model section

@@
 # Model {#sec-model}
+From our analysis of the data, we observed that there were trends between a song's duration, loudness, tempo, and modality for top songs prior 2020 and 2020 onwards.
+We are interested in investigating if a song is likely to be a top song prior to 2020 or not based on what we know about the duration, loudness, tempo, and modality of a top song.
+We will do this by creating a model.
 
+## Model set-up
+In order to predict a top song was a hit prior to 2020, we will make the assumption that there is a relationship between the outcome variable, if the song is a top song prior to 2020 and our variables of interests: duration, loudness, tempo, and modality.
+Our variables of interests will be the predictor variables for the model.
+We will define our model as a logistic regression model with a Bayesian approach as follows:
+\begin{align*} 
+y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{duration}_i + \beta_2 \times \mbox{loudness}_i + \beta_3 \times \mbox{tempo}_i + \beta_4 \times \mbox{modality}_i\\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
+\end{align*}
+
+We define $y_i$ to be a top song, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
+$\pi_i$ is the probability that a top song $i$ is a top song prior to 2020.
+Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB).
+Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
+
+We ran the model using the rstanarm package (@citerstanarm) with the default priors from the package.
+Diagnostics related to the model such as a trace plot, Rhat plot, and posterior predictive check can be found at @sec-appendix.
+
+### Model justification
+A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome.
+However, alternative models were considered such as a multiple linear regression model since the model has multiple predictor variables.
+Other regressions models like negative binomial regressions were also considered but they were not chosen since we are not deleting with discrete counts for the outcome.
+Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounder, lyrics.
+The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020.
+In @fig-causal-model, the reason lyrics is a confounder is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics.
+However, we are unable to observe the effects of lyrics directly with our data.
 ```{r}
@@
 ```
-
-[To Do]
-
-We are interested in investigating if a song is likely to be a top song prior to 2020 or not based on what we know about the duration, loudness, tempo, and modality of a top song.
-
-## Model set-up
-
-We will define our logistic regression model as follows.
-
-\begin{align} 
-y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{duration}_i + \beta_2 \times \mbox{loudness}_i + \beta_3 \times \mbox{tempo}_i + \beta_4 \times \mbox{modality}_i\\
-\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
-\end{align}
-
-We define $y_i$ to be a top song, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
-$\pi_i$ is the probability that a top song $i$ is a top song prior to 2020.
-Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB).
-Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
-
-We ran the model using the rstanarm package of @citerstanarm and we used the default priors from rstanarm.
-
-### Model justification
-[To Do]
 

### 2024-04-16 scripts/00-install_packages.R
Used lintr to fix code styling

@@ -14,16 +14,18 @@ install.packages("janitor") # To clean datasets
 install.packages("dplyr")
 install.packages("ggplot2") # To make graphs
 install.packages("spotifyr") # Spotify API
-install.packages("usethis") 
+install.packages("usethis") # For storing app info for Spotify API
 install.packages("arrow") # To allow Parquet files to be create
 install.packages("ggcorrplot") # For correlation matrix
 install.packages("testthat") # To test
+install.packages("lintr") # To check styling of code
+install.packages("styler") # Styles code
+
 # For models
 install.packages("DiagrammeR")
 install.packages("rsvg")
 install.packages("magrittr")
 install.packages("DiagrammeRsvg")
 install.packages("png")
-
 install.packages("rstanarm")
-install.packages("modelsummary")
+install.packages("modelsummary")# modelsummary package

### 2024-04-16 scripts/00-simulate_data.R

@@ -4,9 +4,8 @@
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R has been ran to install necessary packages.
-
-
+# Pre-requisites: 00-install_packages.R has been ran to install necessary
+# packages.
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
@@ -14,114 +13,119 @@ library(testthat)
 set.seed(646)
 
 #### Simulate data ####
-### Expected Columns: hit_year, track_duration_ms, tempo, loudness, mode_name, key_mode, period, before_pandemic, major, minor
+### Expected Columns: hit_year, track_duration_ms, tempo, loudness, mode_name,
+### key_mode, period, before_pandemic, major, minor
 num_of_top_songs <- 1000
 
-simulated_data <- 
+simulated_data <-
   tibble(
-    hit_year = sample(x=c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023), size = num_of_top_songs, replace = TRUE),
+    hit_year = sample(x = c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021,
+                            2022, 2023), size = num_of_top_songs,
+                      replace = TRUE),
     track_duration_ms = round(runif(num_of_top_songs, min = 0, max = 10000000)),
-    # track_popularity = round(runif(n = num_of_top_songs, min = 0, max = 100)),
-    # energy = round(runif(n = num_of_top_songs, min = 0, max = 1), 1),
+    # (OLD) track_popularity = round(runif(n = num_of_top_songs, min = 0,
+    # max = 100)),
+    # (OLD) energy = round(runif(n = num_of_top_songs, min = 0, max = 1),
+    # 1),
     tempo = runif(n = num_of_top_songs, min = 1, max = 10000000),
     loudness = runif(n = num_of_top_songs, min = -60, max = 0),
-    # valence = round(runif(n = num_of_top_songs, min = 0, max = 1), 1),
-    mode_name = sample(x=c("minor", "major"), size = num_of_top_songs, replace = TRUE),
-    key_mode = sample(x = c("F minor", "F# major", "G# major", "A# minor", "C# minor", "C# major", 
-                            "A major", "C major", "B major", "G major", "D major", "D# major", 
-                            "C minor", "F major", "E minor", "E major", "G minor", "A minor", 
[…]

### 2024-04-16 scripts/01-download_data.R

@@ -4,22 +4,23 @@
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R has been ran to install the necessary packages.
+# Pre-requisites: 00-install_packages.R has been ran to install the necessary
+# packages.
 
 # IMPORTANT NOTE ABOUT THIS SCRIPT:
-# 1. popularity column is updated frequently by Spotify and there is a chance that 
-# the results when replicated might be slightly off from what is described in the paper if 
-# this script is ran. 
+# 1. popularity column is updated frequently by Spotify and there is a chance
+# that the results when replicated might be slightly off from what is described
+# in the paper if this script is ran.
 
-# 2. If this script does not run it is likely because you need to create a 
+# 2. If this script does not run it is likely because you need to create a
 # Spotify Developer Account and a new app for spotifyr. After doing this,
-# call library("usethis") in your console in your RStudio environment, 
-# call edit_r_environ() to open the .Renviron file and add in the following 
-# with the values changed to your new app's Client ID and Client Secret, which can 
-# be found in your new app's settings:
+# call library("usethis") in your console in your RStudio environment,
+# call edit_r_environ() to open the .Renviron file and add in the following
+# with the values changed to your new app's Client ID and Client Secret, which
+# can be found in your new app's settings:
 
-# SPOTIFY_CLIENT_ID = 'PUT_YOUR_CLIENT_ID_HERE'
-# SPOTIFY_CLIENT_SECRET = 'PUT_YOUR_SECRET_HERE'
+# (Spotify Client ID) SPOTIFY_CLIENT_ID = 'PUT_YOUR_CLIENT_ID_HERE'
+# (Spotify Client Secret) SPOTIFY_CLIENT_SECRET = 'PUT_YOUR_SECRET_HERE'
 
 # For more detail, please see https://www.rcharlie.com/spotifyr/
 
@@ -28,8 +29,8 @@ library(tidyverse)
 library(spotifyr)
 
 #### Download data ####
-wickeddreamer96_username <- 'wickeddreamer96'
-antoniomendoza_username <- 'antoniomendoza_'
+wickeddreamer96_username <- "wickeddreamer96"
[…]

### 2024-04-16 scripts/02-data_cleaning.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the dataset on Spotify playlists and outputs a parquet file. 
+# Purpose: Cleans the dataset on Spotify playlists and outputs a parquet file.
 # Author: Emily Su
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R and 01-download_data.R has been ran to install necessary packages
-# and download raw data. 
+# Pre-requisites: 00-install_packages.R and 01-download_data.R has been ran
+# to install necessary packages and download raw data.
 
 #### Workspace setup ####
 library(tidyverse)
@@ -14,23 +14,25 @@ library(arrow)
 library(dplyr)
 
 #### Clean data ####
-# Read in raw data 
-raw_data <- read_csv("data/raw_data/raw_playlists_data.csv", show_col_types = FALSE)
+# Read in raw data
+raw_data <- read_csv("data/raw_data/raw_playlists_data.csv",
+                     show_col_types = FALSE)
 
-# Clean column names 
-cleaned_data <- 
+# Clean column names
+cleaned_data <-
   raw_data |>
   clean_names()
 
-# Select columns: playlist_name, track_name, track_album_name, 
+# Select columns: playlist_name, track_name, track_album_name,
 # track_duration_ms, tempo, loudness, mode_name, key_mode
-cleaned_data <- 
+cleaned_data <-
   cleaned_data |>
-  select(playlist_name, track_name, track_album_name, 
+  select(playlist_name, track_name, track_album_name,
          track_duration_ms, tempo, loudness, mode_name, key_mode)
 
-# Create hit_year, period, before_pandemic, major, and minor columns and remove playlist_name column
-cleaned_data <- 
+# Create hit_year, period, before_pandemic, major, and minor columns
+# and remove playlist_name column
+cleaned_data <-
   cleaned_data |>
   mutate(
     hit_year = case_when(
@@ -63,41 +65,39 @@ cleaned_data <-
     )
   ) |>
   mutate(
-    before_pandemic = case_when(
-      (period == "Before Pandemic") ~ 1,
-      (period == "During and After Pandemic") ~ 0,
-      TRUE ~ 1
-    ) 
-  )|> 
+    before_pandemic = case_when((period == "Before Pandemic") ~ 1,
+                                (period == "During and After Pandemic")
[…]

### 2024-04-16 scripts/03-test_data.R

@@ -1,12 +1,12 @@
 #### Preamble ####
-# Purpose: Tests the clean dataset of top songs. 
+# Purpose: Tests the clean dataset of top songs.
 # Author: Emily Su
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R, 01-download_data.R, and 02-data_cleaning.R 
-# has been ran to install necessary packages, download raw data, and 
-# clean the data, respectively.
+# Pre-requisites: 00-install_packages.R, 01-download_data.R,
+# and 02-data_cleaning.R has been ran to install necessary packages,
+# download raw data, and clean the data, respectively.
 
 #### Workspace setup ####
 library(tidyverse)
@@ -14,85 +14,83 @@ library(arrow)
 library(testthat)
 
 #### Test data ####
-analysis_data <- read_parquet("data/analysis_data/playlists_analysis_data.parquet")
+analysis_data <-
+  read_parquet("data/analysis_data/playlists_analysis_data.parquet")
 
 stopifnot(
   # If all the conditions in here are True, then it will return NULL
-  # 1. hit_year is numeric 
+  # 1. hit_year is numeric
   class(analysis_data$hit_year) == "numeric",
-  # 2. hit_year contains 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023
-  analysis_data$hit_year |> sort() |> unique() == c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023),
-  
-  # 3. Check tempo is numeric 
+  # 2. hit_year contains 2014, 2015, 2016, 2017, 2018, 2019, 2020,
+  # 2021, 2022, 2023
+  analysis_data$hit_year |> sort() |> unique() ==
+    c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023),
+  # 3. Check tempo is numeric
   class(analysis_data$tempo) == "numeric",
   # 4. tempo has a min number that is greater than 0
   analysis_data$tempo |> min() > 0,
   # 5. energy has a max number that is less than or equal to 1.0
-  # analysis_data$energy |> max() <= 1.0,
-  
-  # 6. Check loudness is numeric 
+  # (OLD) analysis_data$energy |> max() <= 1.0,
+  # 6. Check loudness is numeric
   class(analysis_data$loudness) == "numeric",
[…]

### 2024-04-16 scripts/04-model.R

@@ -4,9 +4,10 @@
 # Date: 2 April 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 00-install_packages.R, 01-download_data.R, 02-data_cleaning.R, and
-# 03-test_data.R has been ran to install necessary packages, download raw data, and 
-# clean the data, and test the data, respectively.
+# Pre-requisites: 00-install_packages.R, 01-download_data.R,
+# 02-data_cleaning.R, and 03-test_data.R has been ran to install
+# necessary packages, download raw data, and clean the data,
+# and test the data, respectively.
 
 
 #### Workspace setup ####
@@ -17,14 +18,16 @@ library(arrow)
 set.seed(646)
 
 #### Read data ####
-analysis_data <- read_parquet("data/analysis_data/playlists_analysis_data.parquet")
+analysis_data <-
+  read_parquet("data/analysis_data/playlists_analysis_data.parquet")
 
 ### Model data ####
 
 # Logistic regression model for hit song before 2020/pandemic
 hit_song_model <-
   stan_glm(
-    formula = before_pandemic ~ track_duration_ms + loudness + tempo + mode_name,
+    formula =
+    before_pandemic ~ track_duration_ms + loudness + tempo + mode_name,
     data = analysis_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-04-17 paper/paper.qmd
Updated data section, result section, paper, and install packages script

@@
 hit_song_model_results <- readRDS("../models/hit_song_model.rds")
-
 ```
@@
 These songs often accompany thousands of short clips of people talking about their lives, challenges, etc.
-Ghaffari et al noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
+Ghaffari et al in their study, "The impact of COVID-19 on online music listening behaviors in light of listeners’ social interactions" noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
 We can see this translated through the short-form videos where users don't talk and instead have the song play out to convey how they are feeling or thinking.
@@
 # Data {#sec-data}
-[To Do]
 
-## Variables of Interest
+## Methodology
+The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], testthat [@citetestthat], and lintr [@citelintr].
+The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
 
 ## Data Source and Measurements
+The data used for the model and analysis comes from the ```playlists_analysis_data``` dataset.
+This dataset was constructed by using an R wrapper for Spotify's API called spotifyr [@citespotifyr] to scrape playlists complied by Spotify users antoniomendoza and wickeddreamer96 of songs from the Billboard Year-End Hot 100 singles from 2014 to 2023.
+User antoniomendoza created playlists for the years 2018 to 2023, inclusive, while user wickeddreamer96 created playlists for the years 2014 to 2017, inclusive.
+The latest version of the data used was scraped on Tuesday March 26, 2024 at 9:46pm EDT.
+The Billboard Year-End Hot 100 singles list for the year 2014 onward was created based on data from Nielsen SoundScan [Add citation].
+
+Initially, we had scraped playlists created by Spotify that they have compiled of the top songs on the app.
+However, we had an insufficient amount of observations from these playlists for our dataset since each only had around 50 songs and there were years where Spotify had not created a top song playlist for such as 2021 and 2022.
+
+Each observation in the dataset belongs to a song from the Spotify playlist that we scraped compiled by antoniomendoza and wickeddreamer96 for songs on the Billboard Year-End Hot 100 singles.
+For the duplicate songs that showed up in different years, they were kept as we wanted to analyze song characteristics.
+
+In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals to the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
 ```{r}
@@
 Not seen in @tbl-dataset-preview, before_pandemic, major, and minor are numerical versions of the mode_name and period columns.
+The variables that were originally the raw dataset that was scraped are track_duration_ms, tempo, loudness, mode_name, and key_mode.
 
-## Methodology
-The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], testthat [@citetestthat].
-The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
+The variables constructed from the raw dataset are before_pandemic, period, hit_year, major, and minor.
+How these variables were constructed was that hit_year was obtained from check the name of the playlist and it represents the hit.
+From hit_year, the period column was constructed, which indicates whatever or not the hit year of the song was before 2020, which was when the COVID-19 pandemic started, or during 2020 onwards.
+Reading from the period column, the before_pandemic column was created, where 1 represents that the song was a hit song before the pandemic and 0 otherwise.
+The major and minor columns represents whatever or not the melody of the song is in a major or minor key and the value would be 1 if it's that key and 0 otherwise.
+These columns were constructed from checking mode_name when it would indicate either "major" or "minor".
+
+## Variables of Interest
+As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interests are
 
@@
 
-We ran the model using the rstanarm package (@citerstanarm) with the default priors from the package.
+We ran the model using the rstanarm package [@citerstanarm] with the default priors from the package.
 Diagnostics related to the model such as a trace plot, Rhat plot, and posterior predictive check can be found at @sec-appendix.
@@
 However, we are unable to observe the effects of lyrics directly with our data.
+Weakness and further improvments to the model are discussed in @sec-discussion.
 ```{r}
@@
 ## Model Results
-[To Do]
+@tbl-model-result and @fig-model-result-ci shows the results of our logistic regression model.
+@tbl-model-result indicates that the y-intercept is -2.002, this however has no meaning in this situation since we only have two outcomes, 0 and 1 to represent a hit song from 2020 onwards and a hit song prior to 2020, respectively.
+@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness are constant and its credible interval contains 0.
+[The credible interval means.......] This indicates there is no relationship between a song being a hit before 2020 and the duration of the song.
+Since its credible interval is narrow,
 
@@
 ```
+
 \newpage

### 2024-04-17 scripts/00-install_packages.R

@@ -19,7 +19,6 @@ install.packages("arrow") # To allow Parquet files to be create
 install.packages("ggcorrplot") # For correlation matrix
 install.packages("testthat") # To test
 install.packages("lintr") # To check styling of code
-install.packages("styler") # Styles code
 
 # For models
 install.packages("DiagrammeR")

### 2024-04-17 paper/paper.qmd
Updated model section of paper and updated references

@@
 
-We ran the model using the rstanarm package [@citerstanarm] with the default priors from the package.
+We ran the model using the rstanarm package [@citerstanarm] and R [@citeR] with the default priors from the rstanarm package.
 Diagnostics related to the model such as a trace plot, Rhat plot, and posterior predictive check can be found at @sec-appendix.
@@
 ### Model justification
-A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome.
-However, alternative models were considered such as a multiple linear regression model since the model has multiple predictor variables.
-Other regressions models like negative binomial regressions were also considered but they were not chosen since we are not deleting with discrete counts for the outcome.
-Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounder, lyrics.
-The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020.
-In @fig-causal-model, the reason lyrics is a confounder is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics.
-However, we are unable to observe the effects of lyrics directly with our data.
-Weakness and further improvments to the model are discussed in @sec-discussion.
 ```{r}
@@
 ```
+A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome.
+However, alternative models were considered such as a multiple linear regression model since the model has multiple predictor variables.
+Other regressions models like negative binomial regressions were also considered but they were not chosen since we are not deleting with discrete counts for the outcome.
+Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounding variable, lyrics.
+The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020.
+In @fig-causal-model, the reason lyrics is a confounding variable is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics.
+However, we are unable to observe the effects of lyrics directly with our data.
+Weakness and further improvements to the model are discussed in @sec-discussion.
 
 # Results {#sec-results}
-[To Finish]
+In this section, we visualized our data through graphs and tables as well as present the results from our model.
 
@@
 ### Scale and Modality
-
 ```{r}
@@
 ## Model Results
+
 @tbl-model-result and @fig-model-result-ci shows the results of our logistic regression model.
+@tbl-model-result shows the coefficient estimates obtained for our model with their mean absolute deviation (MAD) while @fig-model-result-ci showcases the coefficient estimates (the dot) alongside their 90% credible interval (the line).
+Coefficient estimates represents....[@citecoefficientsdef].
+Mean absolute deviation is....
+
 @tbl-model-result indicates that the y-intercept is -2.002, this however has no meaning in this situation since we only have two outcomes, 0 and 1 to represent a hit song from 2020 onwards and a hit song prior to 2020, respectively.
-@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness are constant and its credible interval contains 0.
+@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness remains constant and its credible interval.......
 [The credible interval means.......] This indicates there is no relationship between a song being a hit before 2020 and the duration of the song.
-Since its credible interval is narrow,
 
@@
 #| label: tbl-model-result
-#| tbl-cap: How likely a hit song is a hit before 2020 / the pandemic based on its tempo, song duration, loudness, and modality
+#| tbl-cap: Relationships of a hit song is a hit before 2020 / the pandemic with its tempo, song duration, loudness, and modality with their mean absolute deviation (MAD) values in parenthesis. 
 #| echo: false
@@
 ```
+
 ```{r}
 #| label: fig-model-result-ci
-#| fig-cap: The credible intervals for predictors of a hit song before 2020
+#| fig-cap: The credible intervals (line) for coefficient estimates (dot) of predictors of a hit song before 2020
 #| echo: false
@@
 modelplot(hit_song_model_results, conf_level = 0.9) +
-  labs(x = "90 per cent credibility interval")
+  labs(x = "90 per cent credible interval")
 ```

### 2024-04-17 scripts/00-install_packages.R
Added shiny app packages

@@ -19,6 +19,9 @@ install.packages("arrow") # To allow Parquet files to be create
 install.packages("ggcorrplot") # For correlation matrix
 install.packages("testthat") # To test
 install.packages("lintr") # To check styling of code
+# For app
+install.packages("shiny")
+install.packages("shinyWidgets")
 
 # For models
 install.packages("DiagrammeR")

### 2024-04-17 scripts/00-install_packages.R
Added styler to install packages script

@@ -19,6 +19,7 @@ install.packages("arrow") # To allow Parquet files to be create
 install.packages("ggcorrplot") # For correlation matrix
 install.packages("testthat") # To test
 install.packages("lintr") # To check styling of code
+install.packages("styler") # To style code
 # For app
 install.packages("shiny")
 install.packages("shinyWidgets")

### 2024-04-17 other/top-songs-characteristics-app/app.R
Finished shiny app

@@ -1,0 +1,134 @@
+# Purpose: Creates shiny app as a companion to the paper
+# Author: Emily Su
+# Date: 17 April 2024
+# Contact: em.su@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have ran all scripts found in the scripts folder
+
+# Referenced https://mastering-shiny.org/basic-ui.html for code
+
+library(shiny)
+library(shinyWidgets)
+library(ggplot2)
+library(arrow)
+
+# Read in dataset
+playlist_analysis_data <-
+  read_parquet("../../data/analysis_data/playlists_analysis_data.parquet")
+
+# Graph choice
+graph_choice <- c("Track Duration", "Scale and Modality", "Loudness", "Tempo")
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+  # Application title
+  titlePanel(h1("Song Characteristics of Top Songs from Billboard's
+  Year-End Hot 100 singles list (2014 to 2023) before
+             the pandemic and during and after the pandemic",
+    align = "center"
+  )),
+  p('The following are the interactive version of visualizations from the
+      dataset used by the paper "Characteristics of Top Songs Has Changed
+        from Pandemic Brain"', align = "center", width = "50%"),
+
+  # Sidebar with a slider input for number of bins and choose between graphs
+  sidebarLayout(
+    sidebarPanel(
+      p("Created by Emily Su"),
+      p("Last Updated: April 17, 2024"),
+      p("GitHub repository: https://github.com/moonsdust/top-songs"),
+      selectInput(
+        "the_choice", "What song characteristic do you want to
+                  view a graph for?",
+        graph_choice
+      ),
+      sliderInput(
+        inputId = "number_of_bins",
+        label = "Number of bins:",
+        min = 1,
+        max = 50,
+        value = 30
+      ),
+    ),
+
+    # Show a plot based on user's choice
+    mainPanel(plotOutput("curr_plot"))
+  )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+  # Referenced https://stackoverflow.com/questions/38090443/how-to-get-choices
+  # -values-of-selectinput-in-shiny
+  output$curr_plot <- renderPlot({
+    # Scale Plot
+    if (input$the_choice == "Scale and Modality") {
+      proportion_of_each_key_mode <-
+        playlist_analysis_data |>
[…]

### 2024-04-17 other/top-songs-characteristics-app/app.R
Added csv file for shiny app, updated references, shiny app, and data cleaning script

@@ -6,15 +6,16 @@
 # Pre-requisites: Have ran all scripts found in the scripts folder
 
 # Referenced https://mastering-shiny.org/basic-ui.html for code
-
+library(tidyverse)
 library(shiny)
 library(shinyWidgets)
 library(ggplot2)
-library(arrow)
 
 # Read in dataset
-playlist_analysis_data <-
-  read_parquet("../../data/analysis_data/playlists_analysis_data.parquet")
+shiny_app_data <-
+  read_csv("../../data/analysis_data/shiny_app_data.csv",
+    show_col_types = FALSE
+  )
 
 # Graph choice
 graph_choice <- c("Track Duration", "Scale and Modality", "Loudness", "Tempo")
@@ -64,7 +65,7 @@ server <- function(input, output) {
     # Scale Plot
     if (input$the_choice == "Scale and Modality") {
       proportion_of_each_key_mode <-
-        playlist_analysis_data |>
+        shiny_app_data |>
         # Group by if song was a hit before or during and after the pandemic
         group_by(period) |>
         # Count each scale
@@ -90,7 +91,7 @@ server <- function(input, output) {
     }
     # Track Duration Plot
     else if (input$the_choice == "Track Duration") {
-      playlist_analysis_data |>
+      shiny_app_data |>
         ggplot(mapping = aes(x = track_duration_ms)) +
         facet_wrap(facets = vars(period), dir = "v") +
         geom_histogram(
@@ -105,7 +106,7 @@ server <- function(input, output) {
     }
     # Loudness plot
     else if (input$the_choice == "Loudness") {
-      playlist_analysis_data |>
+      shiny_app_data |>
         ggplot(mapping = aes(x = loudness)) +
         facet_wrap(facets = vars(period), dir = "v") +
         geom_histogram(bins = input$number_of_bins, fill = "#ECD078FF") +
@@ -117,7 +118,7 @@ server <- function(input, output) {
     }
     # Tempo plot
     else {
-      playlist_analysis_data |>
+      shiny_app_data |>
         ggplot(mapping = aes(x = tempo)) +
         facet_wrap(facets = vars(period), dir = "v") +
         geom_histogram(bins = input$number_of_bins, fill = "#A7DBD8") +

### 2024-04-17 scripts/02-data_cleaning.R

@@ -99,5 +99,9 @@ cleaned_data <-
   )
 
 #### Save data ####
+# Save as a Parquet file
 write_parquet(cleaned_data,
               "data/analysis_data/playlists_analysis_data.parquet")
+# Save as a CSV file for Shiny App only
+write_csv(cleaned_data,
+          "data/analysis_data/shiny_app_data.csv")

### 2024-04-17 scripts/02-data_cleaning.R
Moved shiny app data

@@ -104,4 +104,4 @@ write_parquet(cleaned_data,
               "data/analysis_data/playlists_analysis_data.parquet")
 # Save as a CSV file for Shiny App only
 write_csv(cleaned_data,
-          "data/analysis_data/shiny_app_data.csv")
+          "other/top-songs-characteristics-app/shiny_app_data.csv")

### 2024-04-17 other/top-songs-characteristics-app/app.R
Finished Shiny app

@@ -13,7 +13,7 @@ library(ggplot2)
 
 # Read in dataset
 shiny_app_data <-
-  read_csv("../../data/analysis_data/shiny_app_data.csv",
+  read_csv("https://raw.githubusercontent.com/moonsdust/top-songs/ec993c0add1e9a2e2c1efa0fe2c3e68623209fed/other/top-songs-characteristics-app/shiny_app_data.csv",
     show_col_types = FALSE
   )
 

### 2024-04-17 paper/paper.qmd
Finished data section, appendix, and finished some of results

@@
 title: "Characteristics of Top Songs Has Changed from Pandemic Brain"
-subtitle: "An analysis of songs on Billboard's Year-End Hot 100 list (2014 to 2023)"
+subtitle: "An analysis of songs on Billboard's Year-End Hot 100 singles list (2014 to 2023)"
 author: Emily Su
@@
 However, this raises the following question, which we explored in our analysis: how are top songs prior to 2020 or the start of the pandemic different from the top songs during and after 2020?
+In our analysis, we defined hit songs before 2020 to be hit songs before the pandemic and hit songs during 2020 and afterwards to be hit songs during and after the pandemic.
 
@@
 In the discussion section (@sec-discussion), we will connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
-Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring an interactive graph of some of the results.
+Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring interactive graphs of some of the results.
 
@@
 ## Methodology
-The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], testthat [@citetestthat], and lintr [@citelintr].
+The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], and testthat [@citetestthat].
+Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].
+Accompanying the paper, a Shiny app [@citeshiny] using shinyWidgets [@citeshinyWidgets] was created containing interactive visualizations.
 The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
@@
 The data used for the model and analysis comes from the ```playlists_analysis_data``` dataset.
-This dataset was constructed by using an R wrapper for Spotify's API called spotifyr [@citespotifyr] to scrape playlists complied by Spotify users antoniomendoza and wickeddreamer96 of songs from the Billboard Year-End Hot 100 singles from 2014 to 2023.
+This dataset was constructed by using an R wrapper for Spotify's API called spotifyr [@citespotifyr] to scrape playlists complied by Spotify users antoniomendoza and wickeddreamer96 of songs from the Billboard Year-End Hot 100 singles from 2014 to 2023 [@citebillboard].
 User antoniomendoza created playlists for the years 2018 to 2023, inclusive, while user wickeddreamer96 created playlists for the years 2014 to 2017, inclusive.
 The latest version of the data used was scraped on Tuesday March 26, 2024 at 9:46pm EDT.
-The Billboard Year-End Hot 100 singles list for the year 2014 onward was created based on data from Nielsen SoundScan [Add citation].
+The Billboard Year-End Hot 100 singles list for the year 2014 onward was created based on data from Nielsen SoundScan [@citenielsen].
+Nielsen SoundScan
+collect data weekly from 39,000 retail outlets, venus, etc.
+in 19 countries such as Canada and the United States on in-store sales and digital sales [@citenielsen].
+They also track how often songs are streamed online and played on the radio [@citenielsen].
 
@@
 For the duplicate songs that showed up in different years, they were kept as we wanted to analyze song characteristics.
-
 In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals to the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
@@
 The variables that were originally the raw dataset that was scraped are track_duration_ms, tempo, loudness, mode_name, and key_mode.
+track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean beats per minutes (BPM) of the song, and loudness is mean loudness of a top song in decibels (dB).
+mode_name represents the key the melody of the song is in, which is either "major" or "minor" and key_mode signifies the scale the song is in.
 
@@
 ## Variables of Interest
-As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interests are
+As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interests are before_pandemic, track_duration_ms, loudness, tempo, and mode_name for our model.
+The major and minor columns were used initially to visualize potential relationships between variables as seen in @fig-correlation-matrix.
+period, hit_year, and key_mode were used to visualize our data prior to training our model and to understand the data.
 
@@
 The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020.
-In @fig-causal-model, the reason lyrics is a confounding variable is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics.
+In @fig-causal-model, the reason lyrics, if there are any, is a confounding variable is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics from the way the lyrics are performed.
 However, we are unable to observe the effects of lyrics directly with our data.
@@
 #| label: fig-mode-bar
-#| fig-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
+#| fig-cap: "Proportion of songs whose modality is in a major or minor key before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
 #| echo: false
@@
   ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
-  facet_wrap(facets = vars(period), dir = "v") + 
-  geom_bar(stat = "identity", binwidth = 3) +
+  facet_wrap(facets = vars(period), dir = "h") + 
+  geom_bar(stat = "identity", binwidth = 3, fill="#e85569") +
   theme_minimal() +
@@
 #| label: tbl-mode-table
-#| tbl-cap: Proportion of songs whose modality is in a major or minor key before 2020 versus during and after 2020
+#| tbl-cap: "Proportion of songs whose modality is in a major or minor key before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
 #| echo: false
@@
 ```
+@fig-mode-bar and @tbl-mode-table reveals that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards).
+Songs in a minor key made up about 40% or less of songs (39% before 2020 and 37% from 2020 onwards) from 2014 to 2023 on Billboard's Year-End Hot 100 singles list.
+There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consist before 2020 and 2020 onwards.
 
@@
 #| label: fig-key-mode-bar
-#| fig-cap: Proportion of songs in different scales before 2020 versus during and after 2020
+#| fig-cap: "Proportion of songs in different scales before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
 #| echo: false
@@
   facet_wrap(facets = vars(period), dir = "v") + 
-  geom_bar(stat="identity", binwidth = 3) +
+  geom_bar(stat="identity", fill="#3E509A") +
   theme_minimal() +
@@
 ```
+When looking into the specific scales of the songs, @fig-key-mode-bar and @tbl-scale-table shows that there was an increase in songs in C major during and after 2020 with C major and C# major made up a majority of songs from Billboard's list from 2020 onwards at approximately 9% each.
+Prior to 2020, the proportion of songs in C major were at 11% however there was a decline in songs in C major after the start of the pandemic.
+There was a decline in songs in G# major from 8% of the list to 6% of the list as well as a an increase in songs in D minor and D# minor from less than 1% of the list to 2% each.
+There were also a decline in songs in F minor where they would make up 6% of Billboard's list prior to 2020 to 4% during and after 2020.
 
@@
 ### Track Duration
-
 ```{r}
 #| label: fig-track-duration-scatterplot
-#| fig-cap: Relationship between the hit year and track duration of a song 
+#| fig-cap: "Relationship between the hit year and track duration of a song on Billboard's Year-End Hot 100 singles.
+Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond.
+Blue curve describes the overall trend."
 #| echo: false
@@
 playlist_analysis_data |> 
-  ggplot(aes(x = hit_year, y = track_duration_ms)) +
+  ggplot(aes(x = hit_year, y = track_duration_ms, color = hit_year)) +
   geom_smooth() +
@@
   labs(x = "Hit year",
-       y = "Duration of song (in ms)")
+       y = "Duration of song (in ms)") + 
+  geom_vline(xintercept = 2020, linetype = "dashed") +
+  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
 ```
@@
 #| label: fig-track-duration-hist
-#| fig-cap: Distribution of track duration (in ms) before 2020 versus during and after 2020  
+#| fig-cap: "Distribution of track duration (in ms) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from songs on Billboard's Year-End Hot 100 singles list."
 #| echo: false
@@
   facet_wrap(facets = vars(period), dir = "v") + 
-  geom_histogram(binwidth = 7000) +
+  geom_histogram(binwidth = 7000, fill="#e85569") +
   theme_minimal() +
@@
 #| label: fig-track-duration-boxplot
-#| fig-cap: Track duration (in ms) before 2020 versus during and after 2020
+#| fig-cap: "Track duration (in ms) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from songs on Billboard's Year-End Hot 100 singles list."
 #| echo: false
@@
 #| label: tbl-track-duration-stats-before
-#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) before 2020
+#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) before 2020 (the pandemic)
 #| echo: false
@@
 #| label: tbl-track-duration-stats-during-after
-#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020
+#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020 (the pandemic)
 #| echo: false
@@
 ```
+@fig-track-duration-scatterplot reveals that songs over the years from 2014 to 2023 decreased slightly in duration overall. @fig-track-duration-hist shows that distribution in
+
+@fig-track-duration-boxplot
 
@@
 #| message: false
-#| 
 playlist_analysis_data |> 
@@
   facet_wrap(facets = vars(period), dir = "v") + 
-  geom_histogram(binwidth = 0.3) +
+  geom_histogram(binwidth = 0.3, fill="#3E509A") +
   theme_minimal() +
@@
   facet_wrap(facets = vars(period), dir = "v") + 
-  geom_histogram(binwidth = 3) +
+  geom_histogram(binwidth = 3, fill="#e85569") +
   theme_minimal() +
@@
 
-## Additional Figures and Tables
+## Shiny application 
+A Shiny application featuring interactive graphs from the paper can be found here:
+https://49z7k8-emily-su.shinyapps.io/top-songs-characteristics-app/
+
+## Posterior predictive check for model {#sec-model-details-posterior-check}
+@fig-posteriorpredictive shows that the posterior distribution of the fitted model is consistent with the actual data. 
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-posteriorpredictive
+#| fig-cap: "How the data impacts how the model fits"
+#| fig-subcap: ["Posterior prediction check for the model"]
+#| layout-ncol: 1
+
+rstanarm::pp_check(hit_song_model_results) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+## Diagnostics for model {#sec-model-details-diagnostics}
+@fig-convergencecheck shows that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
+Both graphs do not show anything abnormal.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheck
+#| fig-cap: "Checking the convergence of the Markov chain Monte Carlo (MCMC) algorithm for hit song model"
+#| fig-subcap: ["Trace plot for model", "Rhat plot for model"]
+#| layout-ncol: 2
+
+plot(hit_song_model_results, "trace")
+
+plot(hit_song_model_results, "rhat") + theme_minimal()
+```
+\newpage
+## Additional Figures
 
@@
 #| label: fig-correlation-matrix
-#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 single and if song was a hit before 2020/the pandemic
+#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 singles list and if song was a hit before 2020/the pandemic
 #| echo: false
@@
 
-## Posterior predictive check for model {#sec-model-details-posterior-check}
+## Additional Tables
+
 ```{r}
+#| label: tbl-scale-table
+#| tbl-cap: "Proportion of songs in different scales before 2020 (the pandemic) versus during and after 2020 (the pandemic) from from Billboard's Year-End Hot 100 singles"
 #| echo: false
-#| eval: true
+#| warning: false
 #| message: false
-#| warning: false
-#| label: fig-posteriorpredictive
-#| fig-cap: "How the data impacts how the model fits"
-#| fig-subcap: ["Posterior prediction check for the model"]
-#| layout-ncol: 1
 
-rstanarm::pp_check(hit_song_model_results) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+proportion_of_each_key_mode |> 
+  rename( 
+    "Period" = period,
+    "Scale" = key_mode,
+    "Count of each scale" = num_of_key_mode,
+    "Proportion of each scale" = proportion_key_mode
+    ) |>
+  kable()
 ```
+
 \newpage
 
-## Diagnostics for model  {#sec-model-details-diagnostics}
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-convergencecheck
-#| fig-cap: "Checking the convergence of the MCMC algorithm for hit song model"
-#| fig-subcap: ["Trace plot for model", "Rhat plot for model"]
-#| layout-ncol: 2
-
-plot(hit_song_model_results, "trace")
-
-plot(hit_song_model_results, "rhat") + theme_minimal()
-```
-
 # References

### 2024-04-17 other/top-songs-characteristics-app/app.R

@@ -5,7 +5,9 @@
 # License: MIT
 # Pre-requisites: Have ran all scripts found in the scripts folder
 
-# Referenced https://mastering-shiny.org/basic-ui.html for code
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
+
 library(tidyverse)
 library(shiny)
 library(shinyWidgets)
@@ -43,6 +45,7 @@ ui <- fluidPage(
                   view a graph for?",
         graph_choice
       ),
+      # Referenced https://mastering-shiny.org/basic-ui.html
       sliderInput(
         inputId = "number_of_bins",
         label = "Number of bins:",

### 2024-04-18 paper/paper.qmd
Finished result section for non-model results and updated model section

@@
 They also track how often songs are streamed online and played on the radio [@citenielsen].
+The list comprise of songs they collected data on starting from December of the previous year to the end of November of the current year, which is when Billboard releases the Year-End Hot 100 list [@citebillboardcalculations].
 
@@
 The variables that were originally the raw dataset that was scraped are track_duration_ms, tempo, loudness, mode_name, and key_mode.
-track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean beats per minutes (BPM) of the song, and loudness is mean loudness of a top song in decibels (dB).
+track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean (average) beats per minutes (BPM) of the song, and loudness is mean loudness of a top song in decibels (dB).
 mode_name represents the key the melody of the song is in, which is either "major" or "minor" and key_mode signifies the scale the song is in.
@@
 #| label: fig-track-duration-scatterplot
-#| fig-cap: "Relationship between the hit year and track duration of a song on Billboard's Year-End Hot 100 singles.
+#| fig-cap: "Relationship between the hit year and track duration of a song (in ms) on Billboard's Year-End Hot 100 singles list.
 Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond.
-Blue curve describes the overall trend."
+Blue, smoother line describes the overall trend and relationship between hit year and track duration."
 #| echo: false
@@
 ```
-
+\newpage
 ```{r}
 #| label: tbl-track-duration-stats-before
-#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) before 2020 (the pandemic)
+#| tbl-cap: "Minimum, quartiles, median, and maximum of track duration (in ms) before 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 #| label: tbl-track-duration-stats-during-after
-#| tbl-cap: Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020 (the pandemic)
+#| tbl-cap: "Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 ```
-@fig-track-duration-scatterplot reveals that songs over the years from 2014 to 2023 decreased slightly in duration overall. @fig-track-duration-hist shows that distribution in
-
-@fig-track-duration-boxplot
+@fig-track-duration-scatterplot reveals that songs over the years from 2014 to 2020 decreased slightly in duration and from 2020 onwards, there was almost no change in trend overall based on the blue, smoother line.
+@fig-track-duration-hist shows that the distribution for top songs before the pandemic and during and after the pandemic are skewed to the right.
+According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic are more disperse around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value, and there are more songs with duration of about 216286 ms.
+@fig-track-duration-boxplot, @tbl-track-duration-stats-during-after, and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length of average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
 
@@
 ### Loudness
-
 ```{r}
 #| label: fig-loudness-scatterplot
-#| fig-cap: Relationship between the hit year and loudness of a song 
+#| fig-cap: "Relationship between the hit year and loudness of a song (in dB) on Billboard's Year-End Hot 100 singles list.
+Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond.
+Blue, smoother line describes the overall trend and relationship between hit year and loudness."
 #| echo: false
@@
 playlist_analysis_data |> 
-  ggplot(aes(x = hit_year, y = loudness)) +
+  ggplot(aes(x = hit_year, y = loudness, color = hit_year)) +
   geom_smooth() +
@@
   labs(x = "Hit year",
-       y = "Loudness of song (in dB)")
+       y = "Loudness of song (in dB)") + 
+  geom_vline(xintercept = 2020, linetype = "dashed") +
+  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
 ```
@@
 #| label: fig-loudness-hist
-#| fig-cap: Distribution of loudness (in dB) before 2020 versus during and after 2020  
+#| fig-cap: "Distribution of loudness (in dB) of hit songs before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 #| label: tbl-loudness-stats-before
-#| tbl-cap: Minimum, quartiles, median, and maximum of loudness of song (in dB) before 2020
+#| tbl-cap: "Minimum, quartiles, median, and maximum of loudness of song (in dB) before 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 #| label: tbl-loudness-stats-during-after
-#| tbl-cap: Minimum, quartiles, median, and maximum of loudness of song (in dB) during and after 2020
+#| tbl-cap: "Minimum, quartiles, median, and maximum of loudness of song (in dB) during and after 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 ```
+@fig-loudness-scatterplot shows that from 2014 to 2023, there was a decrease in loudness overall with hit songs as see with the negative slope of the blue, smoother line.
+@fig-loudness-hist reveals that the distribution of top songs' loudness is left-skewed with the loudness of top songs from during and after the pandemic being more disperse than those before the pandemic around their mean.
+This suggests that top songs from 2020 onwards varied more in terms of their loudness when compared with top songs before 2020.
+@tbl-loudness-stats-before and @tbl-loudness-stats-during-after shows that before the pandemic, top songs on Billboard's list had a mean loudness of -5.964 dB while during and after the pandemic, it had a mean loudness of -6.516 dB.
+This indicates that hit songs during and after the pandemic are less louder than hit songs before the pandemic.
 \newpage
@@
 
+Weakness: There were more songs available prior to 2020 in our dataset. This could have impacted the model.
+
 \newpage

### 2024-04-18 paper/paper.qmd
Added more changes to paper result section

@@
 Nielsen SoundScan
-collect data weekly from 39,000 retail outlets, venus, etc.
+collect data weekly from 39,000 retail outlets, venues, etc.
 in 19 countries such as Canada and the United States on in-store sales and digital sales [@citenielsen].
@@
 Initially, we had scraped playlists created by Spotify that they have compiled of the top songs on the app.
-However, we had an insufficient amount of observations from these playlists for our dataset since each only had around 50 songs and there were years where Spotify had not created a top song playlist for such as 2021 and 2022.
+However, we had an insufficient amount of observations from these playlists for our dataset.
+The playlists we previously were using only had around 50 songs each and there were years where Spotify had not created a top song playlist for such as 2021 and 2022.
 
@@
 A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome.
-However, alternative models were considered such as a multiple linear regression model since the model has multiple predictor variables.
-Other regressions models like negative binomial regressions were also considered but they were not chosen since we are not deleting with discrete counts for the outcome.
+However, alternative models were considered.
+Multiple linear regression model was a candidate for our model since the model has multiple predictor variables but it assumes that the outcome is continuous instead of discrete.
+Other regression models like negative binomial regression were not chosen since we are not dealing with discrete counts for the outcome but instead discrete binary values of 1 and 0.
 Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounding variable, lyrics.
@@
 ```
-\newpage
 ```{r}
@@
 @fig-track-duration-hist shows that the distribution for top songs before the pandemic and during and after the pandemic are skewed to the right.
-According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic are more disperse around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value, and there are more songs with duration of about 216286 ms.
-@fig-track-duration-boxplot, @tbl-track-duration-stats-during-after, and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length of average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
-
-\newpage
+According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic are more disperse around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value.
+Based on the distribution, there are more songs with duration of about 216286 ms for top songs before the pandemic.
+@fig-track-duration-boxplot, @tbl-track-duration-stats-during-after, and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
+The median (the middle value of the data when sorted) track duration also decreased from 214070 ms before 2020 to 195120 ms for 2020 onwards.
 
@@
 ```
-
 ```{r}
@@
 @tbl-loudness-stats-before and @tbl-loudness-stats-during-after shows that before the pandemic, top songs on Billboard's list had a mean loudness of -5.964 dB while during and after the pandemic, it had a mean loudness of -6.516 dB.
+The median loudness of top tracks has also decrease from -5.700 dB prior to the pandemic to -6.151 dB during and after the pandemic.
 This indicates that hit songs during and after the pandemic are less louder than hit songs before the pandemic.
-\newpage
 
 ### Tempo
-
 ```{r}
 #| label: fig-tempo-scatterplot
-#| fig-cap: Relationship between the hit year and tempo of a song 
+#| fig-cap: "Relationship between the hit year and tempo of a song (the beats per minutes (BPM) of a song) on Billboard's Year-End Hot 100 singles list.
+Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond.
+Blue, smoother line describes the overall trend and relationship between hit year and tempo."
 #| echo: false
@@
 playlist_analysis_data |> 
-  ggplot(aes(x = hit_year, y = tempo)) +
+  ggplot(aes(x = hit_year, y = tempo, color = hit_year)) +
   geom_smooth() +
@@
   labs(x = "Hit year",
-       y = "Tempo of song (BPM)")
+       y = "Tempo of song (BPM)") + 
+  geom_vline(xintercept = 2020, linetype = "dashed") +
+  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
 ```
-
 ```{r}
 #| label: fig-tempo-hist
-#| fig-cap: Distribution of tempo (BPM) before 2020 versus during and after 2020  
+#| fig-cap: "Distribution of tempo (BPM) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 ```
-
 ```{r}
 #| label: tbl-tempo-stats-before
-#| tbl-cap: Minimum, quartiles, median, and maximum of tempo of a song (BPM) before 2020
+#| tbl-cap: "Minimum, quartiles, median, and maximum of tempo of a song (BPM) before 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 ```
-
 ```{r}
 #| label: tbl-tempo-stats-during-after
-#| tbl-cap: Minimum, quartiles, median, and maximum of tempo of a song (BPM) during and after 2020
+#| tbl-cap: "Minimum, quartiles, median, and maximum of tempo of a song (BPM) during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
 #| echo: false
@@
 ```
+@fig-tempo-scatterplot reveals almost no change in tempo overall across top songs from 2014 to 2023 based on the blue smoother line appearing to be flat.
+However, looking at @tbl-tempo-stats-before and @tbl-tempo-stats-during-after it shows that both the median and mean tempo of top songs before 2020 (about 120 BPM and 121 BPM, respectively) are lower or about equal than the median and mean tempo of top songs during and after 2020, which is about 120 BPM and 122 BPM, respectively.
+@fig-tempo-hist shows that in both periods of time, their distributions have skewness close to zero however there is less dispersion around the mean for hit songs prior to 2020.
+These results suggest that the the tempo of top songs on Billboard's list are slightly faster during and after 2020.
 
 \newpage
-
 ## Model Results

### 2024-04-18 paper/paper.qmd
Finished initial model results section and further explain model

@@
 
-We define $y_i$ to be a top song, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
-$\pi_i$ is the probability that a top song $i$ is a top song prior to 2020.
+We define $y_i$ to be a top song $i$, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
+$\pi_i$ is the probability of a top song being a hit prior to 2020.
+Following this, $\mbox{logit}(\pi_i)$ is the log odds of a top song $i$ being a top song prior to 2020.
 Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB).
 Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
+$\beta_0$ is the intercept that represents the log odds when all predictor variables are 0.
+$\beta_1$ is the coefficient that signifies the log odds of a song being a hit song prior to 2020 when the $\mbox{duration}_i$ changes by 1 ms.
+$\beta_2$ is the coefficient that stands for the log odds of a song being a hit song prior to 2020 when $\mbox{loudness}_i$ changes by 1 dB.
+$\beta_3$ and $\beta_4$ both represent the log odds of a song being a hit song prior to 2020 when $\mbox{tempo}_i$ and $\mbox{modality}_i$ changes, respectively.
 
@@
 
+\newpage
+
 # Results {#sec-results}
@@
 ```
-@fig-mode-bar and @tbl-mode-table reveals that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards).
-Songs in a minor key made up about 40% or less of songs (39% before 2020 and 37% from 2020 onwards) from 2014 to 2023 on Billboard's Year-End Hot 100 singles list.
-There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consist before 2020 and 2020 onwards.
-
 ```{r}
@@
 ```
+@fig-mode-bar and @tbl-mode-table reveals that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards).
+Songs in a minor key made up about 40% or less of songs (39% before 2020 and 37% from 2020 onwards) from 2014 to 2023 on Billboard's Year-End Hot 100 singles list.
+There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consist before 2020 and 2020 onwards.
+
 When looking into the specific scales of the songs, @fig-key-mode-bar and @tbl-scale-table shows that there was an increase in songs in C major during and after 2020 with C major and C# major made up a majority of songs from Billboard's list from 2020 onwards at approximately 9% each.
@@
 There were also a decline in songs in F minor where they would make up 6% of Billboard's list prior to 2020 to 4% during and after 2020.
-
 \newpage
@@
   geom_histogram(binwidth = 7000, fill="#e85569") +
-  theme_minimal() +
-  labs(x = "Duration of song (in ms)",
-       y = "Count")
-```
-
-```{r}
-#| label: fig-track-duration-boxplot
-#| fig-cap: "Track duration (in ms) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from songs on Billboard's Year-End Hot 100 singles list."
-#| echo: false
-#| warning: false
-#| message: false
-
-playlist_analysis_data |> 
-  ggplot(mapping = aes(x = track_duration_ms)) +
-  facet_wrap(facets = vars(period), dir = "v") + 
-  geom_boxplot() +
   theme_minimal() +
@@
 Based on the distribution, there are more songs with duration of about 216286 ms for top songs before the pandemic.
-@fig-track-duration-boxplot, @tbl-track-duration-stats-during-after, and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
+@tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
 The median (the middle value of the data when sorted) track duration also decreased from 214070 ms before 2020 to 195120 ms for 2020 onwards.
@@
 This indicates that hit songs during and after the pandemic are less louder than hit songs before the pandemic.
+\newpage
 
@@
 @tbl-model-result and @fig-model-result-ci shows the results of our logistic regression model.
-@tbl-model-result shows the coefficient estimates obtained for our model with their mean absolute deviation (MAD) while @fig-model-result-ci showcases the coefficient estimates (the dot) alongside their 90% credible interval (the line).
-Coefficient estimates represents....[@citecoefficientsdef].
-Mean absolute deviation is....
+@tbl-model-result presents the coefficient estimates obtained for our model with their mean absolute deviation (MAD) while @fig-model-result-ci showcases the coefficient estimates (the dot) alongside their 90% credible interval (the line).
+Coefficient estimates represents the change in the log odds of a song being a hit song prior to 2020 when one predictor variable changes by 1 unit as the other predictors stays constant [@citecoefficientsdef].
+With the given data, the 90% credible interval is an interval where the true value of our coefficient estimate has a 90% probability of lying within [@citebayesiandef].
+Mean absolute deviation is the average distance of a data point representing a hit song and the mean.
 
-@tbl-model-result indicates that the y-intercept is -2.002, this however has no meaning in this situation since we only have two outcomes, 0 and 1 to represent a hit song from 2020 onwards and a hit song prior to 2020, respectively.
-@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness remains constant and its credible interval.......
-[The credible interval means.......] This indicates there is no relationship between a song being a hit before 2020 and the duration of the song.
+@tbl-model-result indicates that the intercept $\beta_0$ is -2.002.
+This means that songs in a major key with tempo, loudness, and mode being 0 is less likely to be a top song before 2020.
+@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness remains constant and its 90% credible interval includes the value 0.
+This suggest that when the duration of the song changes the likelihood of a song being a top song prior to 2020 does not change or is impacted.
+When loudness varies and track duration, tempo, and mode stays fixed, @tbl-model-result indicates that its coefficient is 0.148 with a 90% credible interval above 0 as seen in @fig-model-result-ci.
+As the coefficient estimate is positive, it means that hit songs that are louder are more likely to be a hit song prior to 2020.
+Since the coefficient estimate for tempo is −0.002 and its 90% credible interval being slightly below 0 when only tempo varies, it means songs with slower tempos are less likely to be a hit song prior to 2020.
+The coefficient estimate for the modality/mode is 0.095 and its 90% credible interval being wider suggests that top songs that are in a minor key are likely to be a hit before 2020.
+However, the true value for the coefficient for mode can also be negative according to @fig-model-result-ci since the interval passes through 0.
 
@@
 #| label: tbl-model-result
-#| tbl-cap: Relationships of a hit song is a hit before 2020 / the pandemic with its tempo, song duration, loudness, and modality with their mean absolute deviation (MAD) values in parenthesis. 
+#| tbl-cap: Relationships of a hit song being a hit before 2020 / the pandemic with its tempo, song duration, loudness, and modality (mode) with their mean absolute deviation (MAD) values in parenthesis.
 #| echo: false
@@
 
-Weakness: There were more songs available prior to 2020 in our dataset. This could have impacted the model.
+Weakness: There were more songs available prior to 2020 in our dataset. This could have impacted the model and the coefficient estimates seen.
 

### 2024-04-18 paper/paper.qmd
Updated abstract, introduction, and model results

@@
 This paper looks at music characteristics such as tempo, song duration, loudness, and modality of songs from Billboard's Year-End Hot 100 singles list from 2014 to 2023 to reveal patterns and relationships to explain the difference between top songs before 2020 and 2020 onwards.
-The results show that hit songs from 2020 onwards had become on average shorter, quieter, but slightly faster with the melody of songs being in a major key.
+The results show that hit songs from 2020 onwards had become on average quieter but slightly faster with our model revealing that duration does not impact likelihood of a song being a hit prior to 2020 but songs in a minor key were more likely to be a hit before 2020.
 These results can support the evaluation of the emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
@@
 In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020.
-[TODO: ADD RESULTS FROM MODEL].
+In our model, it revealed that top songs prior to 2020 are more likely to be louder, slower, and if its in a minor key, its more likely to be a hit song before 2020.
+However, duration did not impact the likelihood in our model.
 Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight on the overall feelings of different populations during and after the pandemic as well as before it.
@@
 This means that songs in a major key with tempo, loudness, and mode being 0 is less likely to be a top song before 2020.
-@tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness remains constant and its 90% credible interval includes the value 0.
+@tbl-model-result and @fig-model-result-ci reveal that the coefficient estimate of track duration $\beta_1$ is 0.000 when tempo, mode, and loudness remains constant and its 90% credible interval includes the value 0.
 This suggest that when the duration of the song changes the likelihood of a song being a top song prior to 2020 does not change or is impacted.
-When loudness varies and track duration, tempo, and mode stays fixed, @tbl-model-result indicates that its coefficient is 0.148 with a 90% credible interval above 0 as seen in @fig-model-result-ci.
+When loudness varies and track duration, tempo, and mode stays fixed, @tbl-model-result indicates that its coefficient estimate for $\beta_2$ is 0.148 with a 90% credible interval above 0 as seen in @fig-model-result-ci.
 As the coefficient estimate is positive, it means that hit songs that are louder are more likely to be a hit song prior to 2020.
-Since the coefficient estimate for tempo is −0.002 and its 90% credible interval being slightly below 0 when only tempo varies, it means songs with slower tempos are less likely to be a hit song prior to 2020.
-The coefficient estimate for the modality/mode is 0.095 and its 90% credible interval being wider suggests that top songs that are in a minor key are likely to be a hit before 2020.
+Since the coefficient estimate for tempo $\beta_3$ is −0.002 and its 90% credible interval being slightly below 0 when only tempo varies, it means songs with slower tempos are more likely to be a hit song prior to 2020.
+The coefficient estimate for the modality (mode) $\beta_4$ is 0.095 and its 90% credible interval being wider suggests that top songs that are in a minor key are likely to be a hit before 2020.
 However, the true value for the coefficient for mode can also be negative according to @fig-model-result-ci since the interval passes through 0.

### 2024-04-19 scripts/00-simulate_data.R
Added comments in script that lintr was used to check styling

@@ -6,6 +6,7 @@
 # License: MIT
 # Pre-requisites: 00-install_packages.R has been ran to install necessary
 # packages.
+# NOTE: This script was checked through lintr for styling
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)

### 2024-04-19 scripts/01-download_data.R

@@ -6,6 +6,7 @@
 # License: MIT
 # Pre-requisites: 00-install_packages.R has been ran to install the necessary
 # packages.
+# NOTE: This script was checked through lintr for styling
 
 # IMPORTANT NOTE ABOUT THIS SCRIPT:
 # 1. popularity column is updated frequently by Spotify and there is a chance

### 2024-04-19 scripts/02-data_cleaning.R

@@ -6,6 +6,7 @@
 # License: MIT
 # Pre-requisites: 00-install_packages.R and 01-download_data.R has been ran
 # to install necessary packages and download raw data.
+# NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-19 scripts/03-test_data.R

@@ -7,6 +7,7 @@
 # Pre-requisites: 00-install_packages.R, 01-download_data.R,
 # and 02-data_cleaning.R has been ran to install necessary packages,
 # download raw data, and clean the data, respectively.
+# NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-19 scripts/04-model.R

@@ -8,6 +8,7 @@
 # 02-data_cleaning.R, and 03-test_data.R has been ran to install
 # necessary packages, download raw data, and clean the data,
 # and test the data, respectively.
+# NOTE: This script was checked through lintr for styling
 
 
 #### Workspace setup ####

### 2024-04-19 other/top-songs-characteristics-app/app.R
Updated Shiny app to be more interactive

@@ -4,7 +4,7 @@
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Have ran all scripts found in the scripts folder
-
+# NOTE: The code in this script was styled using styler
 # Referenced Telling Stories with Data by Rohan Alexander
 # (https://tellingstorieswithdata.com/) for code
 
@@ -12,6 +12,7 @@ library(tidyverse)
 library(shiny)
 library(shinyWidgets)
 library(ggplot2)
+library(plotly)
 
 # Read in dataset
 shiny_app_data <-
@@ -21,8 +22,10 @@ shiny_app_data <-
 
 # Graph choice
 graph_choice <- c("Track Duration", "Scale and Modality", "Loudness", "Tempo")
+# Colour choice
+colour_choice <- c("#3E509A", "#E85569", "#ECD078FF", "#A7DBD8")
 
-# Define UI for application that draws a histogram
+# Define UI for application
 ui <- fluidPage(
   # Application title
   titlePanel(h1("Song Characteristics of Top Songs from Billboard's
@@ -32,7 +35,9 @@ ui <- fluidPage(
   )),
   p('The following are the interactive version of visualizations from the
       dataset used by the paper "Characteristics of Top Songs Has Changed
-        from Pandemic Brain"', align = "center", width = "50%"),
+        from Pandemic Brain". Toggle over each graph to see more information!',
+    align = "center", width = "50%"
+  ),
 
   # Sidebar with a slider input for number of bins and choose between graphs
   sidebarLayout(
@@ -40,12 +45,17 @@ ui <- fluidPage(
       p("Created by Emily Su"),
       p("Last Updated: April 17, 2024"),
       p("GitHub repository: https://github.com/moonsdust/top-songs"),
+      # Referenced https://mastering-shiny.org/basic-ui.html for code
       selectInput(
         "the_choice", "What song characteristic do you want to
                   view a graph for?",
         graph_choice
       ),
-      # Referenced https://mastering-shiny.org/basic-ui.html
+      selectInput(
+        "colour_choice",
+        "What colour (Hex colour code) do you want the bins to be in?",
+        colour_choice
+      ),
       sliderInput(
         inputId = "number_of_bins",
         label = "Number of bins:",
@@ -56,15 +66,15 @@ ui <- fluidPage(
     ),
 
     # Show a plot based on user's choice
-    mainPanel(plotOutput("curr_plot"))
[…]

### 2024-04-19 scripts/00-install_packages.R

@@ -5,6 +5,7 @@
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: -
+# NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####
 ## Installing packages (only needs to be done once per computer)
@@ -23,6 +24,7 @@ install.packages("styler") # To style code
 # For app
 install.packages("shiny")
 install.packages("shinyWidgets")
+install.packages("plotly")
 
 # For models
 install.packages("DiagrammeR")

### 2024-04-19 scripts/00-simulate_data.R
Removed old tests from scripts

@@ -24,13 +24,8 @@ simulated_data <-
                             2022, 2023), size = num_of_top_songs,
                       replace = TRUE),
     track_duration_ms = round(runif(num_of_top_songs, min = 0, max = 10000000)),
-    # (OLD) track_popularity = round(runif(n = num_of_top_songs, min = 0,
-    # max = 100)),
-    # (OLD) energy = round(runif(n = num_of_top_songs, min = 0, max = 1),
-    # 1),
     tempo = runif(n = num_of_top_songs, min = 1, max = 10000000),
     loudness = runif(n = num_of_top_songs, min = -60, max = 0),
-    # (OLD) valence = round(runif(n = num_of_top_songs, min = 0, max = 1), 1),
     mode_name = sample(x = c("minor", "major"), size = num_of_top_songs, replace
                        = TRUE),
     key_mode = sample(x = c("F minor", "F# major", "G# major", "A# minor",
@@ -56,12 +51,6 @@ stopifnot(
   simulated_data$hit_year |> sort() |> unique() == c(2014, 2015, 2016, 2017,
                                                      2018, 2019, 2020, 2021,
                                                      2022, 2023),
-  # 3. Check that track_popularity is numeric
-  # (OLD) class(simulated_data$track_popularity) == "numeric",
-  # 4. track_popularity has a min number that is equal or greater than 0
-  # (OLD) simulated_data$track_popularity |> min() >= 0,
-  # 5. track_popularity has a max number that is less than or equal to 100
-  # (OLD) simulated_data$track_popularity |> max() <= 100,
   # 6. Check tempo is numeric
   class(simulated_data$tempo) == "numeric",
   # 7. tempo has a min number that is greater than 0
@@ -74,12 +63,6 @@ stopifnot(
   simulated_data$loudness |> min() >= -60,
   # 11. loudness has a max number that is less than or equal to 0
   simulated_data$loudness |> max() <= 0,
-  # 12. Check valence is numeric
-  # (OLD) class(simulated_data$valence) == "numeric",
-  # 13. valence has a min number that is equal or greater than 0.0
-  # (OLD) simulated_data$valence |> min() >= 0.0,
-  # 14. valence has a max number that is less than or equal to 1.0
-  # (OLD) simulated_data$valence |> max() <= 1.0,
   # 15. track_duration_ms is numeric
   class(simulated_data$track_duration_ms) == "numeric",
   # 16. track_duration_ms has a min number that is greater than 0

### 2024-04-19 scripts/03-test_data.R

@@ -38,12 +38,6 @@ stopifnot(
   analysis_data$loudness |> min() >= -60,
   # 8. loudness has a max number that is less than or equal to 0
   analysis_data$loudness |> max() <= 0,
-  # 9. Check valence is numeric
-  # (OLD) class(analysis_data$valence) == "numeric",
-  # 10. valence has a min number that is equal or greater than 0.0
-  # (OLD) analysis_data$valence |> min() >= 0.0,
-  # 11. valence has a max number that is less than or equal to 1.0
-  # (OLD) analysis_data$valence |> max() <= 1.0,
   # 12. track_duration_ms is numeric
   class(analysis_data$track_duration_ms) == "numeric",
   # 13. track_duration_ms has a min number that is greater than 0

### 2024-04-19 paper/paper.qmd
Updated paper

@@
 In the results section (@sec-results), we will reveal tables and graphs made on our datasets, explain what they show, and show our results from our model.
-In the discussion section (@sec-discussion), we will connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
+In the discussion section (@sec-discussion), we will explain what we did, connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
 Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring interactive graphs of some of the results.
@@
 The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], and testthat [@citetestthat].
-Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].
-Accompanying the paper, a Shiny app [@citeshiny] using shinyWidgets [@citeshinyWidgets] was created containing interactive visualizations.
 The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].
@@
 For the duplicate songs that showed up in different years, they were kept as we wanted to analyze song characteristics.
+However, we also obtained duplicated songs from the same year-end playlist when calling the Spotify API and so duplicated songs in this case were removed from the dataset.
 In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals to the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
@@
 
+# Create table displaying data
 playlist_analysis_data |>
@@
 
-# Referenced https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/
-# https://stackoverflow.com/questions/42737645/how-to-export-images-of-diagrammer-in-r
+# Referenced https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/ and https://stackoverflow.com/questions/42737645/how-to-export-images-of-diagrammer-in-r for code
 causal_model <- "digraph {
@@
 }"
-
+# Export the drawing as a png
 exporting_image <- grViz(causal_model) |>
@@
 # Referenced https://stackoverflow.com/questions/23861000/displaying-images-in-r-in-version-3-1-0
+# Read in the image and display it
 the_image <- readPNG("../other/sketches/graph_2_casual_model.png")
@@
 Weakness and further improvements to the model are discussed in @sec-discussion.
-
-\newpage
 
@@
   rename(
-    'num_of_mode_name' = n
+    "num_of_mode_name" = n
   ) |>
@@
 
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 proportion_of_each_mode |>
   ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
+  # Creates two graphs based on each period
   facet_wrap(facets = vars(period), dir = "h") +
-  geom_bar(stat = "identity", binwidth = 3, fill="#e85569") +
+  geom_bar(stat = "identity", binwidth = 3, fill = "#e85569") +
   theme_minimal() +
-  labs(x = "Modality",
-       y = "Proportion") + 
+  labs(
+    x = "Modality",
+    y = "Proportion"
+  ) +
   theme(axis.text.x = element_text(angle = 90))
@@
 
+# Create table for dataframe pertaining to modality
 proportion_of_each_mode |>
@@
   rename(
-    'num_of_key_mode' = n
+    "num_of_key_mode" = n
   ) |>
@@
 
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 proportion_of_each_key_mode |>
   ggplot(mapping = aes(x = key_mode, y = proportion_key_mode)) +
+  # Creates two graphs based on their period
   facet_wrap(facets = vars(period), dir = "v") +
-  geom_bar(stat="identity", fill="#3E509A") +
+  geom_bar(stat = "identity", fill = "#3E509A") +
   theme_minimal() +
-  labs(x = "Scale",
-       y = "Proportion") + 
+  labs(
+    x = "Scale",
+    y = "Proportion"
+  ) +
   theme(axis.text.x = element_text(angle = 90))
@@
   theme_minimal() +
-  labs(x = "Hit year",
-       y = "Duration of song (in ms)") + 
+  labs(
+    x = "Hit year",
+    y = "Duration of song (in ms)"
+  ) +
   geom_vline(xintercept = 2020, linetype = "dashed") +
@@
 #| message: false
-#| 
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 playlist_analysis_data |>
   ggplot(mapping = aes(x = track_duration_ms)) +
+  # Creates two histograms based on period
   facet_wrap(facets = vars(period), dir = "v") +
-  geom_histogram(binwidth = 7000, fill="#e85569") +
+  geom_histogram(binwidth = 7000, fill = "#e85569") +
   theme_minimal() +
-  labs(x = "Duration of song (in ms)",
-       y = "Count")
+  labs(
+    x = "Duration of song (in ms)",
+    y = "Count"
+  )
 ```
@@
   # Rename columns
-  rename (
-    "Duration of song (in ms)"= track_duration_ms
+  rename(
+    "Duration of song (in ms)" = track_duration_ms
   ) |>
@@
 #| message: false
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 
+# Creates scatterplot
 playlist_analysis_data |>
@@
   theme_minimal() +
-  labs(x = "Hit year",
-       y = "Loudness of song (in dB)") + 
+  labs(
+    x = "Hit year",
+    y = "Loudness of song (in dB)"
+  ) +
+  # Creates vertical line to show divide between top songs before 2020 and 2020
+  # onwards
   geom_vline(xintercept = 2020, linetype = "dashed") +
@@
 #| message: false
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 playlist_analysis_data |>
   ggplot(mapping = aes(x = loudness)) +
+  # Creates 2 histograms based on their period
   facet_wrap(facets = vars(period), dir = "v") +
-  geom_histogram(binwidth = 0.3, fill="#3E509A") +
+  geom_histogram(binwidth = 0.3, fill = "#3E509A") +
   theme_minimal() +
-  labs(x = "Loudness of song (in dB)",
-       y = "Count")
+  labs(
+    x = "Loudness of song (in dB)",
+    y = "Count"
+  )
 ```
@@
   # Rename columns
-  rename (
-    "Loudness of song (in dB)"= loudness
+  rename(
+    "Loudness of song (in dB)" = loudness
   ) |>
@@
 #| message: false
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 
+# Creates scatterplot
 playlist_analysis_data |>
   ggplot(aes(x = hit_year, y = tempo, color = hit_year)) +
+  # Creates smoother line
   geom_smooth() +
@@
   theme_minimal() +
-  labs(x = "Hit year",
-       y = "Tempo of song (BPM)") + 
+  labs(
+    x = "Hit year",
+    y = "Tempo of song (BPM)"
+  ) +
+  # Creates vertical line to represent the divide between top songs prior to
+  # 2020 and 2020 onwards
   geom_vline(xintercept = 2020, linetype = "dashed") +
@@
 #| message: false
-
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 playlist_analysis_data |>
   ggplot(mapping = aes(x = tempo)) +
+  # Creates 2 histograms based on their period
   facet_wrap(facets = vars(period), dir = "v") +
-  geom_histogram(binwidth = 3, fill="#e85569") +
+  geom_histogram(binwidth = 3, fill = "#e85569") +
   theme_minimal() +
-  labs(x = "Tempo of song (BPM)",
-       y = "Count")
+  labs(
+    x = "Tempo of song (BPM)",
+    y = "Count"
+  )
 ```
@@
   # Rename columns
-  rename (
-    "Tempo of song (BPM)"= tempo
+  rename(
+    "Tempo of song (BPM)" = tempo
   ) |>
@@
 \newpage
-## Model Results
+## Model Results {#sec-model-results}
 
@@
 modelsummary(
+  # List out information from hit_song_model_results
   list("A hit song before 2020" = hit_song_model_results),
+  # Specify we're using MAD as the statistic
   statistic = "mad"
@@
 #| message: false
-
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
+# Specify credible interval to be 90% and present model results
 modelplot(hit_song_model_results, conf_level = 0.9) +
@@
 # Discussion {#sec-discussion}
+TO REMOVE:
+- What is done in this paper?
+- What is something that we learn about the world?
+What is another thing that we learn about the world?
+(connect back to the real world and explain what the results could mean & explain the implications of our results)
+- What are some weaknesses of what was done? (potential areas of improvement for the paper)
+- What is left to learn or how should we proceed in the future? (suggestions for future works)
 
-[To Do]
+An overview that was done is that we did preliminary analysis of data by graphing relationship between song characteristics and other variables in the dataset such as hit year as seen in @sec-results.
+We also introduced our logistic regression model in a Bayesian framework for top songs prior to 2020 where their results were shown in @sec-model-results.
 
-Weakness: There were more songs available prior to 2020 in our dataset. This could have impacted the model and the coefficient estimates seen.
+## Areas of improvement
+
+There were more songs available prior to 2020 in our dataset. This could have impacted the model and the coefficient estimates seen.
+
+## Next steps
 
@@
 ## Shiny application 
-A Shiny application featuring interactive graphs from the paper can be found here:
-https://49z7k8-emily-su.shinyapps.io/top-songs-characteristics-app/
+Accompanying the paper, a Shiny app [@citeshiny] using shinyWidgets [@citeshinyWidgets] and plotly [@citeplotly] was created that contains interactive visualizations.
+The application can be found here: https://49z7k8-emily-su.shinyapps.io/top-songs-characteristics-app/
+
+## Acknowledgments 
+We would like the acknowledge @citetellingstorieswithdata for some of the R code used in this paper to produce the tables and graphs.
+
+## Note on Reproducing 
+To reproduce the results in the paper, first run the scripts found in the scripts folder of the GitHub repository corresponding to the paper starting with the script, 00-install_packages.R to install the necessary packages.
+
+## Code styling 
+Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].
 
 ## Posterior predictive check for model {#sec-model-details-posterior-check}
-@fig-posteriorpredictive shows that the posterior distribution of the fitted model is consistent with the actual data. 
+@fig-posteriorpredictive shows that the posterior distribution of the fitted model is consistent with the actual data using code modified from @citetellingstorieswithdata. 
 ```{r}
@@
 #| layout-ncol: 1
-
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
 rstanarm::pp_check(hit_song_model_results) +
@@
 @fig-convergencecheck shows that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
-Both graphs do not show anything abnormal.
+Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since the lines in the trace plot fluctuates a lot horizontally and all coefficients are close to 1 and less than 1.1 in the rhat plot.
 ```{r}
@@
 #| layout-ncol: 2
-
-plot(hit_song_model_results, "trace")
-
+# Referenced Telling Stories with Data by Rohan Alexander
+# (https://tellingstorieswithdata.com/) for code
+# Creates trace plot
+plot(hit_song_model_results, "trace") + theme_minimal()
+# Creates rhat plot
 plot(hit_song_model_results, "rhat") + theme_minimal()
@@
   hc.order = TRUE,
-           outline.color = "white", lab = TRUE, legend.title = "Correlation") 
+  outline.color = "white", lab = TRUE, legend.title = "Correlation"
+)
 ```

### 2024-04-19 scripts/01-download_data.R
removed a comment in download data script

@@ -6,12 +6,8 @@
 # License: MIT
 # Pre-requisites: 00-install_packages.R has been ran to install the necessary
 # packages.
-# NOTE: This script was checked through lintr for styling
-
-# IMPORTANT NOTE ABOUT THIS SCRIPT:
-# 1. popularity column is updated frequently by Spotify and there is a chance
-# that the results when replicated might be slightly off from what is described
-# in the paper if this script is ran.
+# IMPORTANT NOTES:
+# 1. This script was checked through lintr for styling
 
 # 2. If this script does not run it is likely because you need to create a
 # Spotify Developer Account and a new app for spotifyr. After doing this,

### 2024-04-19 paper/paper.qmd
Finished paper

@@
 This paper looks at music characteristics such as tempo, song duration, loudness, and modality of songs from Billboard's Year-End Hot 100 singles list from 2014 to 2023 to reveal patterns and relationships to explain the difference between top songs before 2020 and 2020 onwards.
-The results show that hit songs from 2020 onwards had become on average quieter but slightly faster with our model revealing that duration does not impact likelihood of a song being a hit prior to 2020 but songs in a minor key were more likely to be a hit before 2020.
+The results show that hit songs from 2020 onwards had become on average quieter but slightly faster with our model revealing that duration does not impact the likelihood of a song being a hit prior to 2020 but songs in a minor key were more likely to be a hit before 2020.
 These results can support the evaluation of the emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
@@
 From our analysis of the data, we observed that there were trends between a song's duration, loudness, tempo, and modality for top songs prior 2020 and 2020 onwards.
-We are interested in investigating if a song is likely to be a top song prior to 2020 or not based on what we know about the duration, loudness, tempo, and modality of a top song.
+We are interested in investigating if a song is likely to be a top song prior to 2020 on what we know about the duration, loudness, tempo, and modality of a top song.
 We will do this by creating a model.
+Understanding characteristics of songs before 2020 will help us not only understand what made songs hits before the pandemic but also reflect on what makes a song less likely to be a hit before 2020 and instead more likely to be a hit from 2020 onwards.
 
@@
 
-## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic) of songs from the Billboard Year-End Hot 100 singles
+## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic) of songs from Billboard's Year-End Hot 100 singles list
 
@@
 
-### Track Duration
+### Track Duration {#sec-track-duration}
 ```{r}
@@
 
-### Loudness
+### Loudness {#sec-loudness}
 ```{r}
@@
 
-### Tempo
+### Tempo {#sec-tempo}
 ```{r}
@@
 # Discussion {#sec-discussion}
-TO REMOVE:
-- What is done in this paper?
-- What is something that we learn about the world?
-What is another thing that we learn about the world?
-(connect back to the real world and explain what the results could mean & explain the implications of our results)
-- What are some weaknesses of what was done? (potential areas of improvement for the paper)
-- What is left to learn or how should we proceed in the future? (suggestions for future works)
-
-An overview that was done is that we did preliminary analysis of data by graphing relationship between song characteristics and other variables in the dataset such as hit year as seen in @sec-results.
+An overview that was done is that in @sec-results, we did preliminary analysis of the data we obtained from playlists from Spotify by graphing the relationships between song characteristics and other variables in the dataset such as hit year.
 We also introduced our logistic regression model in a Bayesian framework for top songs prior to 2020 where their results were shown in @sec-model-results.
 
+## Songs in a minor key are more likely to be a hit before the pandemic and less so during and after the pandemic 
+From @sec-results, we found that when analyzing the data initially, there were more songs that were in a minor key prior to the pandemic than during it and after it.
+This is also seen in our model result @sec-model-results where it showed that songs in a minor key were more likely a hit before 2020.
+This suggest that songs that are in a minor key are less likely to be a hit during and after the pandemic.
+We also saw from @fig-key-mode-bar there being a decline in songs in different minor scales such as F minor, which made up 6% of Billboard's list before 2020 to 4%.
+There was also an increase in songs in C major with C major and C# major being about equally the most popular scales while before 2020, only songs in C# major were the most popular.
+Music keys can be linked to specific emotions or experiences and it was a common practice to before the 20th century.
+In Christian Schubart's *Ideen zu einer Aesthetik der Tonkunst*, he describes the emotional characteristics of each key [@citescaleemotion].
+When we apply @citescaleemotion's descriptions to songs in the 21st century, C major is described to be simple and innocent and F minor is described to be experiencing a "deep depression" and "longing for the grave".
+@citescaleemotion tends to describe minor keys as more somber while major keys tend to be described as more cheerful.
+Major keys like C major may make people feel nostalgic or joy when listening to the music and therefore there it became one of the most popular key for hit songs to be in after the pandemic.
+From @sec-model-results, @fig-mode-bar, @fig-key-mode-bar, songs in minor keys like F minors are more popular before the pandemic and less so after the start of the pandemic.
+The decline in hit songs being in F minor indicates that people tend to not listen to sadder-sounding songs during the pandemic onwards compared to before 2020.
+These results indicates that people during pandemic had a preference for music that was more cheerful.
+
+## Louder and slower songs are more likely to be hits before the pandemic
+Our model in @sec-model-results reveal that louder and slower songs are more likely to be hits before the pandemic.
+While according to @sec-tempo and @sec-loudness, songs that are quieter and faster were on average hits during and after the pandemic.
+In a study conducted by @citeemotionalresponses, they found that louder music and faster tempo can trigger responses associated with survival responses like increased heart rate.
+This suggest that popular songs from 2014 to 2023 are overall balanced with the emotional response they evoke in listeners in terms of loudness and tempo so that it does not cause too much anxiety in listeners or is too calming.
+
+## Duration of the song does not impact the likelihood of a song being a hit before the pandemic but top songs during and after the pandemic are shorter
+In terms of the duration of the song, our model @sec-model-results suggested that duration of the song does not impact how likely a song being a hit before 2020.
+When looking at @sec-track-duration, top songs after 2019 had on average duration.
+This indicates that prior to 2020, duration of the hit song did not impact it being more likely being a hit song on Billboard's Year-End Hot 100 list.
+However when comparing the track duration of top songs from the two periods, before 2020 and 2020 onwards, top songs are shorter on average during and after 2020.
+During 2020, TikTok experienced its largest number of downloads of around 315 millions download indicating its rise in popularity during the pandemic [@citetiktokstats].
+Videos on the platform are often set to sound clips or snippets from a song and songs that were trending on the platform through dance challenges for example gained popularity outside of TikTok through streaming sites.
+Often catchy songs would gain popularity on TikTok and that the shorter the song is the more replays it would get.
+This could have impacted what songs became popular as well as how long newer songs created after 2019 are.
+However based on our model, prior to 2020, duration did not impact a song's popularity.
+
 ## Areas of improvement
+The analysis was faced with several limitations.
+While creating the dataset, we obtained more songs before 2020 in order to have more observations.
+This is a trade-off for having a more accurate dataset for the model over a dataset that is balanced between the two groups we were observing, which is before the pandemic and during and after the pandemic.
 
-There were more songs available prior to 2020 in our dataset. This could have impacted the model and the coefficient estimates seen.
+Another limitation of our analysis is that the Billboard list itself.
+Since the list also comprises of the year before, for songs on the list that hit their peak popularity from December 2019 to beginning of 2020 before the pandemic started, they might not have representative of the potential changes in terms of song characteristics of songs that become popular after the pandemic started.
+Human error could have occurred when adding music to the playlists we used based on the Billboard Year-End Hot 100 lists where the wrong song could have been added into the playlist for example.
+
+As seen in @fig-causal-model, lyrics is a confounding variable in our model since it can influence the song characteristics in our model and also impact the likelihood of the model's outcome, which was if a top song is a hit before 2020 or not.
+Since we were not able to directly analyze lyrics in our analysis or incorporate it in our model, this could have impacted the model and the coefficient estimates seen.
 
 ## Next steps
+
+Results from the analysis can be used to understand the emotional state of different populations before and after pandemic and assist with how music therapy can be further improved.
+For instance, music therapists could choose or suggests their patients to play songs in specific keys like C major during treatment.
+Further investigations could be conducted on how lyrics impact song characteristics and also how lyrics could impact the popularity of songs.
 
@@
 ```
-
+\newpage
 ## Diagnostics for model {#sec-model-details-diagnostics}

### 2024-04-19 other/top-songs-characteristics-app/app.R

@@ -73,7 +73,9 @@ ui <- fluidPage(
 # Define server logic
 server <- function(input, output) {
   # Referenced https://stackoverflow.com/questions/38090443/how-to-get-choices
-  # -values-of-selectinput-in-shiny
+  # -values-of-selectinput-in-shiny and
+  # https://stackoverflow.com/questions/38917101/how-do-i-show-the-y-value-
+  # on-tooltip-while-hover-in-ggplot2
   output$curr_plot <- renderPlotly({
     # Scale Plot
     if (input$the_choice == "Scale and Modality") {

### 2024-04-19 paper/paper.qmd
Fixed spelling and grammar

@@
 These songs often accompany thousands of short clips of people talking about their lives, challenges, etc.
-Ghaffari et al in their study, "The impact of COVID-19 on online music listening behaviors in light of listeners’ social interactions" noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
+Ghaffari et al in their study, "The impact of COVID-19 on online music listening behaviours in light of listeners’ social interactions" noted that other studies have found that people used music to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping].
 We can see this translated through the short-form videos where users don't talk and instead have the song play out to convey how they are feeling or thinking.
@@
 In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020.
-In our model, it revealed that top songs prior to 2020 are more likely to be louder, slower, and if its in a minor key, its more likely to be a hit song before 2020.
+Our model revealed that top songs prior to 2020 are more likely to be louder, and slower, and if it's in a minor key, it's more likely to be a hit song before 2020.
 However, duration did not impact the likelihood in our model.
-Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight on the overall feelings of different populations during and after the pandemic as well as before it.
+Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight into the overall feelings of different populations during and after the pandemic as well as before it.
 This can aid in improving treatments such as music therapy.
@@
 User antoniomendoza created playlists for the years 2018 to 2023, inclusive, while user wickeddreamer96 created playlists for the years 2014 to 2017, inclusive.
-The latest version of the data used was scraped on Tuesday March 26, 2024 at 9:46pm EDT.
+The latest version of the data used was scraped on Tuesday, March 26, 2024, at 9:46 pm EDT.
 The Billboard Year-End Hot 100 singles list for the year 2014 onward was created based on data from Nielsen SoundScan [@citenielsen].
-Nielsen SoundScan
-collect data weekly from 39,000 retail outlets, venues, etc.
+Nielsen SoundScan collects data weekly from 39,000 retail outlets, venues, etc.
 in 19 countries such as Canada and the United States on in-store sales and digital sales [@citenielsen].
 They also track how often songs are streamed online and played on the radio [@citenielsen].
-The list comprise of songs they collected data on starting from December of the previous year to the end of November of the current year, which is when Billboard releases the Year-End Hot 100 list [@citebillboardcalculations].
+The list comprises of songs they collected data on starting from December of the previous year to the end of November of the current year, which is when Billboard releases the Year-End Hot 100 list [@citebillboardcalculations].
 
 Initially, we had scraped playlists created by Spotify that they have compiled of the top songs on the app.
-However, we had an insufficient amount of observations from these playlists for our dataset.
+However, we had an insufficient number of observations from these playlists for our dataset.
 The playlists we previously were using only had around 50 songs each and there were years where Spotify had not created a top song playlist for such as 2021 and 2022.
@@
 For the duplicate songs that showed up in different years, they were kept as we wanted to analyze song characteristics.
-However, we also obtained duplicated songs from the same year-end playlist when calling the Spotify API and so duplicated songs in this case were removed from the dataset.
-In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals to the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
+However, we also obtained duplicated songs from the same year-end playlist when calling the Spotify API so duplicated songs in this case were removed from the dataset.
+In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
 ```{r}
@@
 The variables that were originally the raw dataset that was scraped are track_duration_ms, tempo, loudness, mode_name, and key_mode.
-track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean (average) beats per minutes (BPM) of the song, and loudness is mean loudness of a top song in decibels (dB).
+track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean (average) beats per minute (BPM) of the song, and loudness is the mean loudness of a top song in decibels (dB).
 mode_name represents the key the melody of the song is in, which is either "major" or "minor" and key_mode signifies the scale the song is in.
@@
 The variables constructed from the raw dataset are before_pandemic, period, hit_year, major, and minor.
-How these variables were constructed was that hit_year was obtained from check the name of the playlist and it represents the hit.
-From hit_year, the period column was constructed, which indicates whatever or not the hit year of the song was before 2020, which was when the COVID-19 pandemic started, or during 2020 onwards.
+How these variables were constructed was that hit_year was obtained by checking the name of the playlist and it represents the hit.
+From hit_year, the period column was constructed, which indicates whether the hit year of the song was before 2020, which was when the COVID-19 pandemic started, or during 2020 onwards.
 Reading from the period column, the before_pandemic column was created, where 1 represents that the song was a hit song before the pandemic and 0 otherwise.
-The major and minor columns represents whatever or not the melody of the song is in a major or minor key and the value would be 1 if it's that key and 0 otherwise.
-These columns were constructed from checking mode_name when it would indicate either "major" or "minor".
+The major and minor columns represent whatever or not the melody of the song is in a major or minor key and the value would be 1 if it's that key and 0 otherwise.
+These columns were constructed by checking mode_name when it would indicate either "major" or "minor".
 
 ## Variables of Interest
-As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interests are before_pandemic, track_duration_ms, loudness, tempo, and mode_name for our model.
+As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interest are before_pandemic, track_duration_ms, loudness, tempo, and mode_name for our model.
 The major and minor columns were used initially to visualize potential relationships between variables as seen in @fig-correlation-matrix.
@@
 # Model {#sec-model}
-From our analysis of the data, we observed that there were trends between a song's duration, loudness, tempo, and modality for top songs prior 2020 and 2020 onwards.
+From our analysis of the data, we observed that there were trends between a song's duration, loudness, tempo, and modality for top songs prior to 2020 and 2020 onwards.
 We are interested in investigating if a song is likely to be a top song prior to 2020 on what we know about the duration, loudness, tempo, and modality of a top song.
 We will do this by creating a model.
-Understanding characteristics of songs before 2020 will help us not only understand what made songs hits before the pandemic but also reflect on what makes a song less likely to be a hit before 2020 and instead more likely to be a hit from 2020 onwards.
+Understanding the characteristics of songs before 2020 will help us not only understand what made songs hits before the pandemic but also reflect on what makes a song less likely to be a hit before 2020 and instead more likely to be a hit from 2020 onwards.
 
 ## Model set-up
-In order to predict a top song was a hit prior to 2020, we will make the assumption that there is a relationship between the outcome variable, if the song is a top song prior to 2020 and our variables of interests: duration, loudness, tempo, and modality.
-Our variables of interests will be the predictor variables for the model.
+To predict whether a top song was a hit before 2020, we will assume that there is a relationship between the outcome variable if the song is a top song before 2020 and our variables of interest: duration, loudness, tempo, and modality.
+Our variables of interest will be the predictor variables for the model.
 We will define our model as a logistic regression model with a Bayesian approach as follows:
@@
 \end{align*}
-
-We define $y_i$ to be a top song $i$, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards.
-$\pi_i$ is the probability of a top song being a hit prior to 2020.
-Following this, $\mbox{logit}(\pi_i)$ is the log odds of a top song $i$ being a top song prior to 2020.
+We define $y_i$ to be a top song $i$, which if it is 1 represents a top song before 2020 and 0 if it is a top song from 2020 onwards.
+$\pi_i$ is the probability of a top song being a hit before 2020.
+Following this, $\mbox{logit}(\pi_i)$ is the log odds of a top song $i$ being a top song before 2020.
 Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB).
-Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
+Set $\mbox{tempo}_i$ to be the mean beats per minute (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.
 $\beta_0$ is the intercept that represents the log odds when all predictor variables are 0.
-$\beta_1$ is the coefficient that signifies the log odds of a song being a hit song prior to 2020 when the $\mbox{duration}_i$ changes by 1 ms.
-$\beta_2$ is the coefficient that stands for the log odds of a song being a hit song prior to 2020 when $\mbox{loudness}_i$ changes by 1 dB.
-$\beta_3$ and $\beta_4$ both represent the log odds of a song being a hit song prior to 2020 when $\mbox{tempo}_i$ and $\mbox{modality}_i$ changes, respectively.
+$\beta_1$ is the coefficient that signifies the log odds of a song being a hit song before 2020 when the $\mbox{duration}_i$ changes by 1 ms.
+$\beta_2$ is the coefficient that stands for the log odds of a song being a hit song before 2020 when $\mbox{loudness}_i$ changes by 1 dB.
+$\beta_3$ and $\beta_4$ both represent the log odds of a song being a hit song before 2020 when $\mbox{tempo}_i$ and $\mbox{modality}_i$ changes, respectively.
 
@@
 ```
-A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome.
+A logistic regression model was chosen because our outcome variable if the song is a top song before 2020, is a binary outcome.
 However, alternative models were considered.
-Multiple linear regression model was a candidate for our model since the model has multiple predictor variables but it assumes that the outcome is continuous instead of discrete.
+The multiple linear regression model was a candidate for our model since the model has multiple predictor variables but it assumes that the outcome is continuous instead of discrete.
 Other regression models like negative binomial regression were not chosen since we are not dealing with discrete counts for the outcome but instead discrete binary values of 1 and 0.
-Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounding variable, lyrics.
-The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020.
-In @fig-causal-model, the reason lyrics, if there are any, is a confounding variable is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics from the way the lyrics are performed.
+Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made before 2020, excluding a confounding variable, lyrics.
+The assumption we make is that there is a relationship between song characteristics and a top song being a hit before 2020.
+In @fig-causal-model, the reason lyrics, if there are any, is a confounding variable is because lyrics can affect not only if the song was a hit before 2020 but also song characteristics from the way the lyrics are performed.
 However, we are unable to observe the effects of lyrics directly with our data.
@@
 ```
-@fig-mode-bar and @tbl-mode-table reveals that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards).
+@fig-mode-bar and @tbl-mode-table reveal that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards).
 Songs in a minor key made up about 40% or less of songs (39% before 2020 and 37% from 2020 onwards) from 2014 to 2023 on Billboard's Year-End Hot 100 singles list.
-There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consist before 2020 and 2020 onwards.
+There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consistent before 2020 and 2020 onwards.
 
-When looking into the specific scales of the songs, @fig-key-mode-bar and @tbl-scale-table shows that there was an increase in songs in C major during and after 2020 with C major and C# major made up a majority of songs from Billboard's list from 2020 onwards at approximately 9% each.
-Prior to 2020, the proportion of songs in C major were at 11% however there was a decline in songs in C major after the start of the pandemic.
-There was a decline in songs in G# major from 8% of the list to 6% of the list as well as a an increase in songs in D minor and D# minor from less than 1% of the list to 2% each.
-There were also a decline in songs in F minor where they would make up 6% of Billboard's list prior to 2020 to 4% during and after 2020.
+When looking into the specific scales of the songs, @fig-key-mode-bar and @tbl-scale-table show that there was an increase in songs in C major during and after 2020 with C major and C# major making up a majority of songs from Billboard's list from 2020 onwards at approximately 9% each.
+Prior to 2020, the proportion of songs in C major was at 11% however there was a decline in songs in C major after the start of the pandemic.
+There was a decline in songs in G# major from 8% of the list to 6% of the list as well as an increase in songs in D minor and D# minor from less than 1% of the list to 2% each.
+There was also a decline in songs in F minor where they would make up 6% of Billboard's list prior to 2020 to 4% during and after 2020.
 \newpage
@@
 @fig-track-duration-scatterplot reveals that songs over the years from 2014 to 2020 decreased slightly in duration and from 2020 onwards, there was almost no change in trend overall based on the blue, smoother line.
-@fig-track-duration-hist shows that the distribution for top songs before the pandemic and during and after the pandemic are skewed to the right.
-According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic are more disperse around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value.
-Based on the distribution, there are more songs with duration of about 216286 ms for top songs before the pandemic.
-@tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms.
+@fig-track-duration-hist shows that the distribution of top songs before the pandemic and during and after the pandemic are skewed to the right.
+According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic is more dispersed around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value.
+Based on the distribution, there are more songs with a duration of about 216286 ms for top songs before the pandemic.
+@tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the pandemic with a mean track duration of 216286 ms.
 The median (the middle value of the data when sorted) track duration also decreased from 214070 ms before 2020 to 195120 ms for 2020 onwards.
@@
 ```
-@fig-loudness-scatterplot shows that from 2014 to 2023, there was a decrease in loudness overall with hit songs as see with the negative slope of the blue, smoother line.
-@fig-loudness-hist reveals that the distribution of top songs' loudness is left-skewed with the loudness of top songs from during and after the pandemic being more disperse than those before the pandemic around their mean.
+@fig-loudness-scatterplot shows that from 2014 to 2023, there was a decrease in loudness overall with hit songs as seen with the negative slope of the blue, smoother line.
+@fig-loudness-hist reveals that the distribution of top songs' loudness is left-skewed with the loudness of top songs from during and after the pandemic being more dispersed than those before the pandemic around their mean.
 This suggests that top songs from 2020 onwards varied more in terms of their loudness when compared with top songs before 2020.
-@tbl-loudness-stats-before and @tbl-loudness-stats-during-after shows that before the pandemic, top songs on Billboard's list had a mean loudness of -5.964 dB while during and after the pandemic, it had a mean loudness of -6.516 dB.
-The median loudness of top tracks has also decrease from -5.700 dB prior to the pandemic to -6.151 dB during and after the pandemic.
-This indicates that hit songs during and after the pandemic are less louder than hit songs before the pandemic.
+@tbl-loudness-stats-before and @tbl-loudness-stats-during-after show that before the pandemic, top songs on Billboard's list had a mean loudness of -5.964 dB while during and after the pandemic, they had a mean loudness of -6.516 dB.
+The median loudness of top tracks has also decreased from -5.700 dB prior to the pandemic to -6.151 dB during and after the pandemic.
+This indicates that hit songs during and after the pandemic are louder than hit songs before the pandemic.
 \newpage
@@
 @fig-tempo-scatterplot reveals almost no change in tempo overall across top songs from 2014 to 2023 based on the blue smoother line appearing to be flat.
-However, looking at @tbl-tempo-stats-before and @tbl-tempo-stats-during-after it shows that both the median and mean tempo of top songs before 2020 (about 120 BPM and 121 BPM, respectively) are lower or about equal than the median and mean tempo of top songs during and after 2020, which is about 120 BPM and 122 BPM, respectively.
+However, looking at @tbl-tempo-stats-before and @tbl-tempo-stats-during-after shows that both the median and mean tempo of top songs before 2020 (about 120 BPM and 121 BPM, respectively) are lower or about equal to the median and mean tempo of top songs during and after 2020, which is about 120 BPM and 122 BPM, respectively.
 @fig-tempo-hist shows that in both periods of time, their distributions have skewness close to zero however there is less dispersion around the mean for hit songs prior to 2020.
-These results suggest that the the tempo of top songs on Billboard's list are slightly faster during and after 2020.
+These results suggest that the tempos of top songs on Billboard's list are slightly faster during and after 2020.
 
@@
 
-@tbl-model-result and @fig-model-result-ci shows the results of our logistic regression model.
+@tbl-model-result and @fig-model-result-ci show the results of our logistic regression model.
 @tbl-model-result presents the coefficient estimates obtained for our model with their mean absolute deviation (MAD) while @fig-model-result-ci showcases the coefficient estimates (the dot) alongside their 90% credible interval (the line).
-Coefficient estimates represents the change in the log odds of a song being a hit song prior to 2020 when one predictor variable changes by 1 unit as the other predictors stays constant [@citecoefficientsdef].
+A coefficient estimate represents the change in the log odds of a song being a hit song prior to 2020 when one predictor variable changes by 1 unit as the other predictors stay constant [@citecoefficientsdef].
 With the given data, the 90% credible interval is an interval where the true value of our coefficient estimate has a 90% probability of lying within [@citebayesiandef].
-Mean absolute deviation is the average distance of a data point representing a hit song and the mean.
+The mean absolute deviation is the average distance of a data point representing a hit song and the mean.
 
 @tbl-model-result indicates that the intercept $\beta_0$ is -2.002.
-This means that songs in a major key with tempo, loudness, and mode being 0 is less likely to be a top song before 2020.
-@tbl-model-result and @fig-model-result-ci reveal that the coefficient estimate of track duration $\beta_1$ is 0.000 when tempo, mode, and loudness remains constant and its 90% credible interval includes the value 0.
-This suggest that when the duration of the song changes the likelihood of a song being a top song prior to 2020 does not change or is impacted.
-When loudness varies and track duration, tempo, and mode stays fixed, @tbl-model-result indicates that its coefficient estimate for $\beta_2$ is 0.148 with a 90% credible interval above 0 as seen in @fig-model-result-ci.
+This means that songs in a major key with tempo, loudness, and mode being 0 are less likely to be top songs before 2020.
+@tbl-model-result and @fig-model-result-ci reveal that the coefficient estimate of track duration $\beta_1$ is 0.000 when tempo, mode, and loudness remain constant and its 90% credible interval includes the value 0.
+This suggests that when the duration of the song changes the likelihood of a song being a top song prior to 2020 does not change or is impacted.
+When loudness varies and track duration, tempo, and mode stay fixed, @tbl-model-result indicates that its coefficient estimate for $\beta_2$ is 0.148 with a 90% credible interval above 0 as seen in @fig-model-result-ci.
 As the coefficient estimate is positive, it means that hit songs that are louder are more likely to be a hit song prior to 2020.
-Since the coefficient estimate for tempo $\beta_3$ is −0.002 and its 90% credible interval being slightly below 0 when only tempo varies, it means songs with slower tempos are more likely to be a hit song prior to 2020.
+Since the coefficient estimate for tempo $\beta_3$ is −0.002 and its 90% credible interval is slightly below 0 when only tempo varies, it means songs with slower tempos are more likely to be a hit song prior to 2020.
 The coefficient estimate for the modality (mode) $\beta_4$ is 0.095 and its 90% credible interval being wider suggests that top songs that are in a minor key are likely to be a hit before 2020.
@@
 # Discussion {#sec-discussion}
-An overview that was done is that in @sec-results, we did preliminary analysis of the data we obtained from playlists from Spotify by graphing the relationships between song characteristics and other variables in the dataset such as hit year.
+An overview that was done is that in @sec-results, we did a preliminary analysis of the data we obtained from playlists from Spotify by graphing the relationships between song characteristics and other variables in the dataset such as hit year.
 We also introduced our logistic regression model in a Bayesian framework for top songs prior to 2020 where their results were shown in @sec-model-results.
@@
 ## Songs in a minor key are more likely to be a hit before the pandemic and less so during and after the pandemic 
-From @sec-results, we found that when analyzing the data initially, there were more songs that were in a minor key prior to the pandemic than during it and after it.
+From @sec-results, we found that when analyzing the data initially, there were more songs that were in a minor key prior to the pandemic than during and after it.
 This is also seen in our model result @sec-model-results where it showed that songs in a minor key were more likely a hit before 2020.
-This suggest that songs that are in a minor key are less likely to be a hit during and after the pandemic.
-We also saw from @fig-key-mode-bar there being a decline in songs in different minor scales such as F minor, which made up 6% of Billboard's list before 2020 to 4%.
+This suggests that songs that are in a minor key are less likely to be a hit during and after the pandemic.
+We also saw from @fig-key-mode-bar that there was a decline in songs in different minor scales such as F minor, which made up 6% of Billboard's list before 2020 to 4%.
 There was also an increase in songs in C major with C major and C# major being about equally the most popular scales while before 2020, only songs in C# major were the most popular.
-Music keys can be linked to specific emotions or experiences and it was a common practice to before the 20th century.
+Music keys can be linked to specific emotions or experiences and it was a common practice before the 20th century.
 In Christian Schubart's *Ideen zu einer Aesthetik der Tonkunst*, he describes the emotional characteristics of each key [@citescaleemotion].
@@
 @citescaleemotion tends to describe minor keys as more somber while major keys tend to be described as more cheerful.
-Major keys like C major may make people feel nostalgic or joy when listening to the music and therefore there it became one of the most popular key for hit songs to be in after the pandemic.
-From @sec-model-results, @fig-mode-bar, @fig-key-mode-bar, songs in minor keys like F minors are more popular before the pandemic and less so after the start of the pandemic.
+Major keys like C major may make people feel nostalgic or joyful when listening to the music and therefore there it became one of the most popular keys for hit songs to be in after the pandemic.
+From @sec-model-results, @fig-mode-bar, @fig-key-mode-bar, songs in minor keys like F minors were more popular before the pandemic and less so after the start of the pandemic.
 The decline in hit songs being in F minor indicates that people tend to not listen to sadder-sounding songs during the pandemic onwards compared to before 2020.
-These results indicates that people during pandemic had a preference for music that was more cheerful.
+These results indicate that people during the pandemic had a preference for music that was more cheerful.
 
 ## Louder and slower songs are more likely to be hits before the pandemic
-Our model in @sec-model-results reveal that louder and slower songs are more likely to be hits before the pandemic.
-While according to @sec-tempo and @sec-loudness, songs that are quieter and faster were on average hits during and after the pandemic.
+Our model in @sec-model-results reveals that louder and slower songs are more likely to be hits before the pandemic.
+According to @sec-tempo and @sec-loudness, songs that are quieter and faster were on average hits during and after the pandemic.
 In a study conducted by @citeemotionalresponses, they found that louder music and faster tempo can trigger responses associated with survival responses like increased heart rate.
-This suggest that popular songs from 2014 to 2023 are overall balanced with the emotional response they evoke in listeners in terms of loudness and tempo so that it does not cause too much anxiety in listeners or is too calming.
+This suggests that popular songs from 2014 to 2023 are overall balanced with the emotional response they evoke in listeners in terms of loudness and tempo so that it does not cause too much anxiety in listeners or are too calming.
 
 ## Duration of the song does not impact the likelihood of a song being a hit before the pandemic but top songs during and after the pandemic are shorter
-In terms of the duration of the song, our model @sec-model-results suggested that duration of the song does not impact how likely a song being a hit before 2020.
-When looking at @sec-track-duration, top songs after 2019 had on average duration.
-This indicates that prior to 2020, duration of the hit song did not impact it being more likely being a hit song on Billboard's Year-End Hot 100 list.
-However when comparing the track duration of top songs from the two periods, before 2020 and 2020 onwards, top songs are shorter on average during and after 2020.
-During 2020, TikTok experienced its largest number of downloads of around 315 millions download indicating its rise in popularity during the pandemic [@citetiktokstats].
+In terms of the duration of the song, our model @sec-model-results suggested that the duration of the song does not impact how likely a song is to be a hit before 2020.
+When looking at @sec-track-duration, top songs after 2019 had on average a shorter duration.
+This indicates that prior to 2020, the duration of the hit song did not impact it being more likely being a hit song on Billboard's Year-End Hot 100 list.
+However, when comparing the track duration of top songs from the two periods, before 2020 and 2020 onwards, top songs are shorter on average during and after 2020.
+During 2020, TikTok experienced its largest number of downloads of around 315 million downloads indicating its rise in popularity during the pandemic [@citetiktokstats].
 Videos on the platform are often set to sound clips or snippets from a song and songs that were trending on the platform through dance challenges for example gained popularity outside of TikTok through streaming sites.
-Often catchy songs would gain popularity on TikTok and that the shorter the song is the more replays it would get.
+Often catchy songs would gain popularity on TikTok and the shorter the song is the more replays it would get.
 This could have impacted what songs became popular as well as how long newer songs created after 2019 are.
@@
 Another limitation of our analysis is that the Billboard list itself.
-Since the list also comprises of the year before, for songs on the list that hit their peak popularity from December 2019 to beginning of 2020 before the pandemic started, they might not have representative of the potential changes in terms of song characteristics of songs that become popular after the pandemic started.
-Human error could have occurred when adding music to the playlists we used based on the Billboard Year-End Hot 100 lists where the wrong song could have been added into the playlist for example.
+Since the list also considers the year before, songs on the list that hit their peak popularity from December 2019 to the beginning of 2020 before the pandemic started might not be representative of the potential changes in terms of song characteristics of songs that became popular after the pandemic started.
+Human error could have occurred when adding music to the playlists we used based on the Billboard Year-End Hot 100 lists where the wrong song could have been added to the playlist for example.
 
@@
 
-Results from the analysis can be used to understand the emotional state of different populations before and after pandemic and assist with how music therapy can be further improved.
-For instance, music therapists could choose or suggests their patients to play songs in specific keys like C major during treatment.
+Results from the analysis can be used to understand the emotional state of different populations before and after the pandemic and assist with how music therapy can be further improved.
+For instance, music therapists could choose or suggest their patients to play songs in specific keys like C major during treatment.
 Further investigations could be conducted on how lyrics impact song characteristics and also how lyrics could impact the popularity of songs.

### 2024-04-19 paper/paper.qmd
Updated wording

@@
 In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020.
-Our model revealed that top songs prior to 2020 are more likely to be louder, and slower, and if it's in a minor key, it's more likely to be a hit song before 2020.
+Our model revealed that top songs before 2020 are more likely to be louder, and slower, and if it is in a minor key, it is more likely to be a hit song before 2020.
 However, duration did not impact the likelihood in our model.
@@
 In the discussion section (@sec-discussion), we will explain what we did, connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works.
-Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring interactive graphs of some of the results.
+Finally, the appendix section (@sec-appendix) will contain extra tables and graphs from our results, additional information about the model, a link to a Shiny application featuring interactive graphs of some of the results, and other information.
 
@@
 Reading from the period column, the before_pandemic column was created, where 1 represents that the song was a hit song before the pandemic and 0 otherwise.
-The major and minor columns represent whatever or not the melody of the song is in a major or minor key and the value would be 1 if it's that key and 0 otherwise.
+The major and minor columns represent whatever or not the melody of the song is in a major or minor key and the value would be 1 if it is that key and 0 otherwise.
 These columns were constructed by checking mode_name when it would indicate either "major" or "minor".

### 2024-04-19 paper/paper.qmd
Fixed causal model in paper and fixed formatting

@@
 
-### Model justification
+## Model justification
 ```{r}
 #| label: fig-causal-model
-#| fig-cap: Causal relationship between song characteristics and hit song (before 2020 or during and after 2020)
+#| fig-cap: Causal relationship between song characteristics and hit song before 2020
 #| echo: false
@@
   node [shape = plaintext]
-    A [label = 'Hit song before 2020 or during and after 2020']
+    A [label = 'Hit song before 2020']
     B [label = 'Modality']
